Wetness modulates the effects of grazing on net ecosystem productivity in global grasslands

Version 1

Abstract

Overgrazing has caused widespread global concern. However, how grazing intensity affects ecosystem carbon dioxide (CO2) fluxes with wetness fluctuations in grasslands at a global scale remains poorly understood. Here we measured ecosystem CO2 fluxes covering the 7th to 11th years of a continuous grazing experiment in a typical steppe, and conducted a meta-analysis of 585 observations by collecting relevant data in global grasslands. Our experimental results showed that mean gross primary productivity (GPP) and ecosystem respiration (ER) were decreased under heavy grazing (HG) compared with control without grazing (CK), whereas net ecosystem productivity (NEP) was not different between HG and CK in the typical steppe. The NEP was higher under light grazing in the wetter years, but it was not different between LG and CK under drier years in the typical steppe. In contrast, grazing decreased GPP, ER and NEP, respectively, in global grasslands. Both our field experiment and the meta-analysis revealed that the response of NEP to light grazing was linearly correlated with the wetness index, which may be mediated by greater plant relative growth rate in wetter years. Higher wetness and aboveground biomass (AGB) increased the response of NEP to grazing in global grasslands. However, heavy grazing reduced NEP and AGB even under higher wetness indices in global grasslands, resulting from a loss of their resilience in long-term heavy grazing. These findings indicate that light grazing appears to be a promising management to promote plant relative growth rate and CO2 flux sequestration. Furthermore, major drivers influencing ecosystem CO2 fluxes in response to grazing were aboveground biomass, grazing intensity and wetness index. Overall, this study provides a global perspective on the response of ecosystem CO2 fluxes to grazing intensity and improves our knowledge of the factors influencing the response of ecosystem CO2 fluxes to grazing intensity.

Introduction

Grasslands play a vital role in livestock production, promoting carbon (C) sequestration and regulating climate change (Heimann & Reichstein, 2008; Zhou et al., 2017). Globally, grassland ecosystems cover approximately one fifth of the world land surface and contain about 20-30% of the global pool of soil organic C (Du et al., 2022). Grazing is a widespread land-use activity that significantly alters terrestrial C dynamics (He et al., 2020; Zhou et al., 2025). Up to 49.25% of grassland area has degraded worldwide, around 5% experiencing severe degradation (Gang et al., 2014). Grassland degradation is influenced by a multitude of driving factors, such as overgrazing and precipitation limitation (Sloat et al., 2018). Furthermore, the C balance of grasslands is also increasingly influenced by changes in grazing management practices and water availability (Hilker et al., 2014; Lemoine et al., 2016; Ren et al., 2025). Overgrazing and drier conditions could potentially drive grasslands from acting as C sinks to C sources (Hao et al., 2018; Morgan et al., 2016; Shao et al., 2013).

Gross primary productivity (GPP) and ecosystem respiration (ER) are critical processes governing C sequestration and emission in terrestrial ecosystems (Chen et al., 2023; Knauer et al., 2023). Net ecosystem productivity (NEP) is the balance between ecosystem C uptake and release, and serves as a key indicator of the net C exchange (Niu et al., 2010; Zhang et al., 2024). Grazing influences NEP by altering GPP and ER (Fig. 1). Thus, accurately assessing ecosystem carbon dioxide (CO2) fluxes (GPP, ER and NEP) in grasslands is essential to the terrestrial C cycle at local, regional and global scales. However, current evidence on the effect of grazing on ecosystem CO2 fluxes in grasslands remains inconsistent, with studies reporting both positive (Chang et al., 2020; Du et al., 2022) and negative (Liu et al., 2020; Wang et al., 2023; Zhu et al., 2015) effects. For instance, some studies demonstrated that grazing decreased GPP and ER (Nakano et al., 2020; Qi et al., 2024; Wang et al., 2025), while others reported an increase or no difference in GPP or ER (Chang et al., 2020; Luo et al., 2015; Shi et al., 2022a). Similarly, some studies investigated that grazing significantly reduced NEP (Wang et al., 2023; Zhu et al., 2015), while others reported an increase or no difference in NEP under grazing (Du et al., 2022; Gomez-Casanovas et al., 2018; Yu et al., 2025). The variations in the responses of ecosystem CO2 fluxes are likely driven by differences in grazing intensities, climatic variations, livestock and grassland types (Hou et al., 2016; Wang et al., 2023; Xiao et al., 2011; Zhu et al., 2015).

Conceptual illustration of how grazing with wetness fluctuation would affect net ecosystem productivity (A), and geographic (B) and climatic distributions of data (C) included in this meta-analysis.

The effects of grazing intensities on ecosystem CO2 fluxes in global grasslands were also varied in different studies (Table S1; Jiang et al., 2020; Rong et al., 2017; Zhang et al., 2022). Specifically, heavy grazing usually reduced the GPP, ER and NEP (Ju et al., 2024; Rong et al., 2017; Zhang et al., 2022), but a meta-analysis also reported the response of NEP was increased in Chinese grasslands (Jiang et al., 2020). Generally, light grazing can enhance NEP by promoting plant regeneration and improving photosynthetic efficiency, whereas moderate to heavy grazing tends to reduce NEP, largely as a result of decreased vegetation coverage and plant biomass (Yu et al., 2025). Conversely, a meta-analysis showed that the response of NEP was decreased under light and moderate grazing (Zhang et al., 2022). However, it was reported that moderate grazing led to the highest NEP (Zhu et al., 2015). These conflicting results underscore the complex responses of ecosystem CO2 fluxes in grasslands to grazing and highlight the necessity to identify key factors controlling ecosystem CO2 flux dynamics under grazing management. This understanding is essential for accurately modelling and predicting grassland C dynamics under global change.

Many studies have demonstrated that yearly variation in grassland NEP can be substantial, primarily driven by fluctuating climatic conditions, as they can directly regulate vegetation growth via altering soil moisture and temperature (Rong et al., 2017; Wayne Polley et al., 2008; Shen et al., 2016). Annual precipitation is one of the climatic parameters, while the wetness index (WI) serves as a more integrative climatic indicator that incorporates both precipitation and temperature, thereby reflecting the overall water surplus or deficit (Song et al., 2019). A higher wetness index (WI > 30) indicates sufficient water availability for plant growth, whereas a lower wetness index (WI≤ 30) suggests the water availability may be limited (Botzan et al., 1998). Considerable uncertainty remains regarding how NEP responds to grazing under different water conditions. Although evidence suggested that, under wet year, grazing significantly increased NEP; under dry year, grazing suppressed NEP and can even turn NEP negative (Shao et al., 2013). In general, grazing tends to have a less suppressive effect on NEP under wet than dry conditions (Okach et al., 2019; Zhou et al., 2019). For example, heavy grazing consistently reduced NEP, with a more pronounced decline occurring during dry years in a desert steppe (Jin et al., 2023). In addition, overgrazing led to decoupling of precipitation patterns and ecosystem carbon exchange in the desert steppe, primarily by altering plant community composition (Wang et al., 2023). However, it remains unclear how NEP responds to grazing intensities under water availability fluctuations across global grasslands.

Previous studies on the effects of grazing intensities on ecosystem CO2 fluxes have been constrained by limited spatial and temporal scales, which has led to an incomplete understanding of how different grazing intensities influence ecosystem CO2 fluxes with changes of annual wetness index. In this study, we investigated the effects of grazing intensities on ecosystem CO2 fluxes by combining a long-term (7-11 years) field experiment conducted in a typical steppe and a meta-analysis of global grasslands. The objectives of this study were to: (i) investigate the effects of grazing intensity with annual wetness fluctuations on ecosystem CO2 fluxes (GPP, ER and NEP) covering the 7th to 11th years of a continuous grazing experiment in a typical steppe as well as the meta-analysis in global grasslands; (ii) explore how environmental factors (particularly wetness index, soil moisture and temperature, grazing duration and grazing intensity) regulate the effects of grazing on ecosystem CO2 fluxes. We hypothesized that: (H1) the response of net ecosystem productivity to grazing (including light, moderate and heavy grazing) would shift from negative to positive with increasing wetness index; (H2) aboveground biomass (AGB), grazing intensity and wetness index would act as key factors regulating the responses of ecosystem CO2 fluxes to grazing, with their responses intensifying with AGB, grazing intensity and wetness index.

Materials and methods

Experimental site

The grazing experiment was conducted at Xilinhot National Climate Observatory (44°12′ N, 116°19′ E, 1129 m above sea level), in the middle part of Inner Mongolian typical steppe in China. The average annual precipitation was 291 mm from 2019 to 2023, with average precipitation in the growing season (May to September) varying from 168 to 325 mm (Fig. S1). The average mean air temperature was 3.9 °C (from 2019 to 2023). The soil is classified as a Haplic Calcisol based on the FAO soil classification system or a Calcic-Orthic Aridisol based on the USDA soil classification system (Liang et al., 2021). The vegetation community is dominated by Stipa grandis P. Smirn (perennial bunchgrass) and Leymus chinensis Trin. Tzvel (perennial rhizomatous grass) (Wu et al., 2024).

Experimental design

The grazing experiment was initiated in 2013 and was based on a randomized design (Fig. S2). There were four treatments with three replicates: control without grazing (CK), light grazing (LG, 0.75 sheep ha-1 year-1), moderate grazing (MG, 1.5 sheep ha-1 year-1) and heavy grazing (HG, 3 sheep ha-1 year-1) conducted in plots with 1.44 ha (120 × 120 m2) established in the research station (Fig. S2). Two-year-old sheep similar in size were selected to use for grazing from May to September in each year. Sheep were grazing from 7:00 AM to 6:00 PM each day, and then gathered into an enclosure at night (Liang et al., 2021).

Measurement of ecosystem CO2 fluxes

Ecosystem CO2 fluxes were measured using a transparent chamber (0.5 × 0.5 × 0.8 m3) connected to an infrared gas analyzer (LI-8100, Lincoln, NE, USA). To ensure an airtight seal during measurements, the junction between the chamber bottom and the aluminum base was secured with tapes. After positioning the chamber, the enclosed air was continuously mixed by two fans in the chamber. The CO2 concentrations were recorded 12 times at 10-second intervals over a 2-minute period. The NEP was then determined from the rate of CO2 concentration change, combined with chamber volume and headspace temperature data. A positive NEP value indicated net ecosystem CO2 uptake, whereas a negative value denoted net ecosystem CO2 release. Following each NEP measurement, the chamber was removed to vent the headspace and then covered with a dark cloth to inhibit photosynthesis before being repositioned on the same base for ER measurement. Both the NEP and ER were measured throughout the growing seasons from 2019 to 2023, corresponding to the 7th to 11th years of grazing. The GPP was thus calculated as the sum of NEP and ER using data collected over this five-year period. All gas exchange measurements were conducted on sunny days between 9:00 and 11:00 AM, with a frequency of three times per month.

Climate conditions and soil microclimate

The mean annual precipitation (MAP) and mean annual temperature (MAT) for the study area from 2019 to 2023 were obtained from the China Meteorological Data Service Center (http://data.cma.cn/). The De Martonne wetness index (WI) was calculated as follows (Song et al., 2019): (1) Wetness index(WI)=MAPMAT+10 Soil temperature and moisture (hereafter referred to as soil moisture) at a depth of 0-10 cm were measured using the probes of the LI-8100 system. Measurement sites for soil temperature and moisture were selected near the collars randomly placed in the plots. They were investigated synchronized with the measurements of ecosystem CO2 fluxes.

Plant sampling and analyses

To track changes in plant AGB, vegetation surveys were conducted in August each year from 2019 to 2023. Specifically, in each plot, five 1 × 1 m2 quadrats were randomly placed. All aboveground biomass within these quadrats was clipped at ground level and then oven-dried at 65 °C for 48 h to determine the community-level dry biomass. Belowground biomass (BGB) was quantified by collecting root biomass using soil cores. Each September from 2019 to 2023, two soil cores with a diameter of 7 cm were collected from each 1 × 1 m2 quadrat at depths of 0-30 cm. Subsequently, the belowground parts were extracted from soil by rinsing them in water, and then oven-dried at 65 °C for 48 h to obtain the dry weight.

The relative growth rate (RGR) was defined as the ratio of aboveground biomass regeneration to the number of days between two consecutive grazing events and was calculated by Eq. (2) (Zong & Shi, 2019): (2) RGR=(AGBEi+1−AGBi)Δt where AGBEi+1: the aboveground biomass inside the caged plots after the (i+1)-th grazing round, AGBi: the remaining aboveground biomass after the i-th grazing, △t: the number of days between the i-th and (i+1)-th grazing rounds.

Aboveground net primary productivity (ANPP) was estimated as the sum of the total herbage consumed by livestock throughout the growing season and the AGB measured at the end of the final grazing rotation of the season.

Data Analysis

All the analyses were conducted using the R (version 4.4.3) (R Core Team, 2023). Annual mean values were calculated for GPP, ER, NEP, AGB, BGB, root shoot ratio (R/S), soil temperature, and soil moisture from 2019 to 2023. Differences across years and among treatments within each year were examined using the Least Significant Difference (LSD) test. Response ratios of AGB, BGB, NEP, and GPP were calculated based on Eq. (3). Additionally, linear regression analysis was conducted using the “lm” function to evaluate the relationships of GPP, NEP, AGB, BGB, and RGR with WI. (3) ln⁡RR=ln⁡(XtXC)=ln⁡(Xt)−ln⁡(XC) where Xt and XC are the mean values of the abovementioned indicators for the grazing (t) and control (c) treatments, respectively.

The repeated measures ANOVA (the ezANOVA function of the “ez” R package) was used to examine the effects of grazing treatment and sampling years on ecosystem CO2 fluxes (GPP, ER, and NEP), microclimatic variables (soil temperature and moisture), and plant biomass parameters (AGB, BGB, and R/S) across the five-year study period from 2019 to 2023.

We also conducted a random forest model to identify the significant environmental predictors of GPP, ER, NEP, AGB and BGB. The random forest model was implemented in R using the “randomForest” and “rfPermute” packages. In this model, the importance of each predictor was evaluated based on the percentage increase in the mean squared error (MSE) when that predictor was permuted; a higher MSE value indicated greater importance of the factor. The MSE values for each decision tree, based on out-of-bag estimates from the random forest model, were generated using the “rfPermute” package to assess the relative importance of the predictor variables (Liao et al., 2024).

Meta-analysis of grazing experiments in global grasslands

To assess the effects of grazing on GPP, ER, and NEP in global grasslands, we conducted a systematic literature search for relevant papers published up to June 8, 2025, using the Web of Science (http://apps.webofknowledge.com/) and China National Knowledge Infrastructure (CNKI, https://www.cnki.net). The keywords and terms used were (“grazing” OR “stocking” OR “livestock” OR “grazing intensity”) AND (“net ecosystem productivity” OR “net ecosystem exchange” OR “gross ecosystem productivity” OR “ecosystem respiration” OR “soil carbon flux” OR “soil CO2” OR “soil carbon flux” or “soil carbon emission”). Relevant studies were selected based on the following criteria: (a) Studies were conducted using simulating grazing experiments, and studies with trampling and mowing were excluded. (b) The control and grazing plots were under similar conditions, such as dominant species, community composition, soil type and climate. (c) Studies involving additional treatments (e.g., fertilization, experimental warming, or precipitation manipulation) were excluded, and only the control and grazing treatments were included in our data. (d) The mean and sample size for any measurements of GPP, ER, and NEP under both grazing and control conditions must be reported. Finally, 62 published papers with 585 observations of GPP, ER, and NEP responses (n = 161, 249, and 175, respectively) were selected (Fig. 4), and the geographical location of these observations were presented in Fig. 1. Meanwhile, the study site, elevation, MAP, MAT, and grazing intensity from the selected papers were recorded. The wetness indices based on MAP and MAT of each site were calculated according to Eq. (1). Where available, data for soil temperature, soil moisture, AGB, BGB, and R/S were also recorded.

In addition, to better explore the effects of grazing intensity and environmental factors on ecosystem CO2 fluxes, the data was classified into different subgroups by grazing intensity (LG, MG, and HG), mean annual temperature (≤5 °C and >5 °C), wetness index (≤ 30 and > 30), and precipitation (≤ 400 mm and > 400 mm).

The effect size of grazing on each variable was quantified using the natural logarithm of response ratios (lnRR) based on Eq. (3) (Feng et al., 2023).

The weighing of the ith study (WRR) was calculated as (Pittelkow et al., 2015): (4) WRR=(NC×NT)(NC+NT) where NT and NC are the replicate numbers of treatment and control, respectively.

A linear mixed-effects model, with ‘study’ included as a random factor, was employed to estimate the weighted response ratio (RR++) across studies or within a specific group. (5) RR=β0+πstudy+ε where β0 is the coefficient, πstudy denotes the random effect associated with ‘study’ (accounting for autocorrelation among observations from the same study), and ɛ corresponds to the residual sampling error. When the assumption of normality was violated, bootstrapping with 999 iterations was performed using the BOOT package to derive the 95% confidence interval (CI) for each RR++ (Chen et al., 2021). The percentage change (%) of the response ratio was calculated as follows: (6) Percentage Change=(eRR++−1)×100% The lnRR++ was considered statistically significant if its 95% confidence interval (CI) did not overlap zero (Chen et al., 2021).

The random forest model was conducted using “randomForest” and “rfPermute” packages to identify the relative importance of the predictor variables (wetness index, grazing duration and intensity, the responses of soil moisture and temperature and livestock types) of GPP, ER, NEP, AGB and BGB in the meta-analysis (Liao et al., 2024).

Results

Grazing intensity effects on ecosystem CO2 fluxes and plant biomass

Continuous heavy grazing reduced the mean GPP by 27.31% and the mean ER by 28.05% in the grazing experiment; however, NEP did not differ between HG and CK (Fig. 2). The mean GPP and ER were decreased by 17.15% and 19.48%, respectively, in MG than in CK, but NEP was not different between MG and CK. The mean ER was decreased by 19.48% and 28.05%, respectively, in MG and HG than in LG. The mean NEP was decreased by 43.55% in HG than in LG, but was not different among LG, MG and CK. The GPP and NEP were 29.27% and 85.83% higher, respectively, in LG than in CK, but the ER was not different in 2020, the year with the highest precipitation and WI among the five years (Fig. S1 and 2). The mean aboveground and belowground biomass was decreased by 34.96% and 19.44%, respectively, in HG than in CK, but the mean root shoot ratio was increased by 30.57% in HG than in CK (Fig. 2; P < 0.05).

Effects of different grazing intensities on ecosystem carbon dioxide (CO2) fluxes and biomass in our field experiment.

CK, enclosure; LG, light grazing; MG, moderate grazing; HG, heavy grazing. Data (means ± SE, n = 3) followed by different lowercase letters indicate differences at P < 0.05 between treatments.

Wetness was the most important predictor of NEP, while grazing intensity was the most important predictor of GPP and ER, revealing a strong coupling between CO2-driven changes in wetness index and grazing intensity (Fig. S3). Grazing intensity and duration were also the important predictors influencing AGB and BGB. The linear regression analysis showed that overall grazing as well as light grazing stimulated the responses of GPP and NEP with increasing wetness index (Fig. 3; P < 0.05). This relationship could be explained by the significant positive correlation between NEP and both ANPP and RGR (Fig. S4). In the wettest year of 2020, the relative growth rate peaked relative to other years, reaching levels 3.16, 2.76 and 1.96 times higher under LG, MG, and HG respectively, compared to CK (Fig. S5). The NEP increased with WI under HG (Fig. S6; P < 0.01), whereas the response ratio of NEP to heavy grazing was not significantly correlated with WI (Fig. 3B).

Relationships of the responses of gross primary productivity (GPP) and net ecosystem productivity (NEP) with potential regulating factors from 2019 to 2023 under different grazing intensity.

Overall, overall grazing; LG, light grazing; MG, moderate grazing; HG, heavy grazing. The significant regression lines and 95% confidence intervals are shown with lines (solid for significant) and shaded areas, respectively.

Synthesis of grazing on ecosystem CO2 fluxes and plant biomass across global grasslands

In the meta-analysis across global grasslands, grazing significantly decreased ecosystem CO2 fluxes by 13.56% to 30.46% (Fig. 4). Significant differences were observed in the responses of ecosystem CO2 fluxes to different grazing intensities across global grasslands (Fig. 5A). Heavy grazing had a more pronounced negative effect on ecosystem CO2 fluxes compared to light and moderate grazing. The response of GPP was decreased by 15.89%, 11.88% and 26.34%, respectively, under LG, MG and HG. The response of ER was decreased by 7.42%, 14.91% and 21.82%, respectively, under LG, MG and HG. The response of NEP was decreased by 18.86% and 34.89%, respectively, in LG and HG, but was not significant in MG. Light, moderate and heavy grazing resulted in reductions of AGB by 27.89%, 34.57% and 66.17% (Fig. 5B). Moderate and heavy grazing decreased the BGB by 14.85% and 26.13%, respectively, but the response of BGB was not significant in LG. Grazing increased the root shoot ratio by 31.78% to 111.11%. Heavy grazing had a more pronounced positive effect on root shoot ratio compared to light and moderate grazing.

Effects of grazing on gross primary productivity, ecosystem respiration, net ecosystem productivity, aboveground biomass, belowground biomass and root shoot ratio, soil moisture and soil temperature, in global grasslands in the field experiment in this study (A) and from the meta-analysis (B).

Circles and error bars represent average parameter estimates and 95% confidence interval (CI). The star (*) indicates significance when the CI did not overlap zero. Sample sizes of observations for each variable are displayed on the left.

Effects of different grazing intensities on ecosystem CO2 fluxes, biomass and soil temperature and moisture in global grasslands from the meta-analysis.

Circles and error bars represent average effects and their 95% confidence interval (CI). The star (*) indicates significant effects, i.e., the 95% CI did not overlap zero. The sample size for each variable is shown on the left. LG, light grazing; MG, moderate grazing; HG, heavy grazing.

Globally, grazing decreased GPP, ER and NEP under low precipitation conditions (≤ 400 mm), while the reductions were not significant under high precipitation conditions (> 400 mm) (Fig. S7A). Grazing significantly reduced GPP, ER, and NEP at mean annual temperatures ≤ 5 °C, but these effects were not significant at temperatures > 5 °C (Fig. S7B). Similarly, under lower wetness conditions (WI ≤ 30), grazing decreased GPP, ER, and NEP more than under higher wetness conditions (WI > 30) (Fig. 6). The response of NEP was decreased under lower WI (WI ≤ 30) in MG and HG, but it was not different under higher WI (WI > 30) in MG. The response of NEP was decreased under higher WI (WI > 30) in HG. Higher wetness index mitigated the response of NEP to grazing. The duration of grazing also influenced its effect: GPP and ER declined more substantially under longer grazing durations (Fig. S7C). The most pronounced decline in NEP occurred under grazing durations of 5 to 10 years (P < 0.001).

Responses of ecosystem CO2 fluxes and biomass to grazing (A) and different grazing intensities (B-C), at wetness index ≤ 30 and wetness index > 30 in global grasslands from the meta-analysis.

Circles and error bars represent average parameter estimates and 95% confidence interval (CI). The sample size for each variable is shown on the left. LG, light grazing; MG, moderate grazing; HG, heavy grazing.

In global grasslands, the key environmental factors influencing GPP and ER response to grazing included WI, grazing duration, the response of soil moisture, and grazing intensity (Fig. S8; P < 0.05). For NEP, the major predictors were WI, grazing duration and intensity (P < 0.05). The most important environmental parameters for AGB response to grazing were grazing intensity, WI, grazing duration, the response of soil moisture and temperature and livestock types. For BGB, the major predictors were WI, grazing duration, the response of soil moisture and temperature. The response of NEP was positively correlated with WI under light, moderate and total grazing in global grasslands (Fig. 7; P<0.1). The response of ecosystem CO2 fluxes was affected by AGB under grazing in global grasslands (Fig. S9).

Relationships of the responses of ecosystem CO2 fluxes with potential regulating factors in global grasslands from the meta-analysis.

The marginal (m) and conditional (c) R2 indicate the proportion of variance explained by the fixed effects and by both the fixed and random effects study), respectively. The black solid lines and gray shaded areas in panels are the mean and 95% confidence interval of the slope, respectively. The sizes of points are proportional to their corresponding weights. RR, response ratio; overall, total grazing; LG, light grazing; MG, moderate grazing; HG, heavy grazing.

Discussion

The responses of ecosystem CO2 fluxes and plant biomass to grazing and controlling factors

Grazing decreased the responses of ecosystem CO2 fluxes and plant biomass in global grasslands, but only NEP was not significantly affected by grazing in our field experiment (Fig. 4 and S10). The lack of a significant response in NEP may be because the site in this field study was managed for year-round continuous low-intensity grazing (Liang et al., 2021). Grazing stimulated the NEP under wet conditions, but depressed it under dry conditions, a pattern consistent across the typical steppe in our experiment and supported by our global meta-analysis (Figs. 3B and 7H). This finding aligned with our first hypothesis (H1) that the response of NEP increased with the increase in wetness index under grazing. This relationship could be explained by the significant positive correlation of NEP with AGB (Fig. S9). The linear regression analysis also showed that light grazing stimulated the responses of NEP with increasing wetness index. This relationship could be explained by the significant positive correlation between NEP and both ANPP and RGR (Fig. S4). Light grazing usually stimulates leaf regrowth following defoliation, and these new leaves often are more physiologically active than the older leaves that contribute much of leaf area on ungrazed treatment (Wayne Polley et al., 2008), which likely imply a stronger leaf photosynthesis and C sink (Reich et al., 2007). Under lower wetness conditions (WI ≤ 30), light grazing reduced GPP and ER to a lesser extent than moderate and heavy grazing (Fig. 6B). Compared with higher grazing intensities, light grazing consumed less forage by livestock, thereby better mitigated the impact of lower moisture on GPP and ER. Moderate grazing decreased the response of GPP and NEP under lower WI, but the response of GPP and NEP were not different under moderate grazing of higher WI in global grasslands (Figs. 6 and 8), indicating that higher wetness offset the response of GPP and NEP to moderate grazing. This was likely because grazing-induced biomass reduction was less under wet condition in grasslands (Jin et al., 2023). However, not all grazing intensities aligned with the first hypothesis, likely due to severe loss of vegetation resilience under heavy grazing. Heavy grazing, even under higher wetness, reduced NEP and AGB in global grasslands (Fig. 2, 5 and 6C), resulting in a severe decline or even loss of their resilience, which could not be restored to control levels in long-term heavy grazing (Irob et al., 2023).

Schematic summary of the effects of grazing intensity on ecosystem carbon dioxide (CO2) fluxes and biomass in global grasslands in the meta-analysis.

The arrows indicate negative effects. Asterisks (*) indicate significant effects on variables at P < 0.05. GPP, gross primary productivity; ER, ecosystem respiration; NEP, net primary productivity; LG, light grazing; MG, moderate grazing; HG, heavy grazing.

Both our experiment and meta-analysis showed that aboveground biomass, grazing intensity and wetness index were key factors regulating the responses of ecosystem CO2 fluxes to grazing (Fig. S3, S8 and S9), with their responses intensifying with higher aboveground biomass and wetness index, which aligned with our second hypothesis (H2). Grazing reduced aboveground biomass and plant respiration, thereby decreasing GPP and ER (Fig. 4; Fig. S10; Liu et al., 2020; Wang et al., 2025). Plants enhanced photosynthetic structures and physiological functions under grazing pressure. These adaptations improved CO2 assimilation and allocate more carbohydrates to compensate for the loss of aboveground biomass, thereby achieving compensatory growth (Wang et al, 2020; Yu et al., 2025). The increase in GPP and NEP in LG in 2020 was due to the increased relative growth rate in LG than in CK (Fig. S5), suggesting that grazing stimulated greater plant compensatory growth in years with more precipitation (Li et al., 2025). Moreover, appropriate grazing intensity facilitated the removal of senescent plant tissues, leading to increased leaf light capture capability and enhanced community photosynthetic capacity (Shen et al., 2019). The aboveground and belowground biomass, as well as root shoot ratio of plants were influenced by grazing activities such as feeding, trampling, and excretion (Cao et al., 2024). In this study, plant biomass (both aboveground and belowground biomass) was lower under heavy grazing. Under heavy grazing conditions, where soil nutrients were often depleted, plants tended to prioritize biomass allocation to roots to enhance nutrient acquisition and utilization (Liu et al., 2024). Moreover, heavy grazing can stimulate C allocation to roots (Ren et al., 2024), which explains why the root/shoot ratio was higher under heavy grazing (Fig. 2). In the wetter years (2020 and 2021), the aboveground biomass was higher while the root shoot ratio was lower than in the drier years. Heavy grazing affected the allocation strategy of aboveground and belowground biomass by plants.

Limitations and implications for future study

Our results highlight that light grazing serves as a promising management strategy to promote aboveground net primary productivity and CO2 sequestration. However, some aspects still need to be further explored. Our study has limitations due to the relatively small sample size, which stems from the scarcity of global-scale studies exploring the effects of grazing intensity on ecosystem CO2 fluxes. The data of ecosystem CO2 fluxes from large-scale eddy covariance flux towers were not included in this meta-analysis, because there is an absence of a robust model to convert data between flux tower observations and chamber-based methods (Shi et al., 2022b). In addition, future studies could conduct more experiments of ecosystem CO2 fluxes in South American, Africa and Oceania, which have large amounts of grasslands for grazing.

Regardless of these limitations, our study has significant implications for future study and grassland production. The primary factors regulating ecosystem CO2 flux responses to grazing were identified as aboveground biomass, wetness index and grazing intensity. The NEP was positively affected by wetness index in LG, MG and HG. However, in both the typical steppe grazing experiment and the synthesis of global grasslands, the NEP resilience increased with higher wetness index specifically under light grazing, but not under heavy grazing. During wetter years, the NEP was even higher under LG compared to CK, whereas in drier years, no significant difference was observed between LG and CK (Fig. 2). This pattern aligns with the greater RGR and ANPP observed under elevated wetness index in light grazing treatment (Fig. S4). Globally, higher wetness index and aboveground biomass enhanced the response of NEP to grazing (Figs. 3, 7 and S9). However, even under higher wetness conditions, heavy grazing reduced NEP and AGB in global grasslands (Fig. 6), reflecting a severe decline, or even loss of resilience under long-term heavy grazing pressure. These results indicate that the negative impacts of long-term heavy grazing on NEP could not be offset by higher annual wetness index in global grasslands, underscoring the necessity of implementing strict grazing management measures. Additionally, moderate grazing under wetter years may be an appropriate management strategy to sustain grassland carbon balance while supporting livestock production.

In summary, this study presents the first comprehensive assessment of how annual wetness index affecting the response of ecosystem CO2 fluxes to grazing across global grasslands. Our results showed that wetness index and aboveground biomass were key factors regulating the responses of ecosystem CO2 fluxes, and both exhibiting positive associations with these responses. Grazing reduced the response of GPP and ER in our field experiment of typical steppe and in the meta-analysis of global grasslands. Higher wetness indices offset the effect of moderate grazing rather than heavy grazing on NEP in global grasslands. These findings underscore the importance of developing region-specific grazing strategies guided by local wetness indices.

eLife Assessment

This study uses convincing modeling methods and analyses of rich behavioral datasets to investigate the role of attention in value-based decision making; for instance, as when choosing between two snacks. The results are valuable, as they challenge existing theories that assume that paying attention to an available option biases the eventual choice toward that option. The results suggest that the correlation between attention and decision-making is formed largely after and not before the (internal) choice process has terminated, a finding that offers an intuitively appealing rethinking of how attention and decision-making processes interact during value-based choices.

Reviewer #1 (Public review):

Summary:

This study examines whether gaze direction actively shapes choice during food preference decisions or whether gaze and choice evolve largely independently until the moment of commitment. The established framework in this context, the aDDM, assumes that gaze causally biases the accumulation of evidence in favour of the fixated item. The authors show convincingly that this model fails to fit key behavioural patterns across several datasets, as do other published models that make the same assumption. The authors propose an alternative model (Post-Decision-Gaze or PDG) in which gaze and decision formation are decoupled: gaze does not influence the decision process, nor is it drawn toward the ultimately chosen item, until after the decision threshold is reached. Only during the motor execution period (after commitment) is gaze directed to the chosen option. They demonstrate that this model fits several observed patterns better than the aDDM and related variants.

Strengths:

The work thoroughly considers multiple models and datasets. It advances an interesting alternative perspective on gaze-decision interactions and highlights meaningful shortcomings in existing models. The authors take the time to explain how modelling assumptions produce specific patterns in the data, which is certainly insightful to readers interested in the modelling of value-based decision making.

Weaknesses:

It is unclear to what extent the model's success relies on the way non-decision time is formalised in the model. In the proposed PDG model, non-decision time is decomposed into separate visual encoding, saccadic execution, and manual execution components. Several values (assumed or recovered) do not match known physiological or behavioural ranges. This is a common issue in the literature, and the authors may want to address it in light of broader work discussing what non-decision time consists of in both manual and saccadic actions (e.g., Bompas et al., 2024, Non decision time: the Higgs boson of decision, Psychological Review).

In particular, the "saccadic execution" parameter appears far too long and too variable to reflect merely execution; instead, it likely includes decisional components. This would make more sense since manual and saccadic planning essentially rely on distinct brain areas, hence it seems unrealistic that crossing a single threshold would trigger both manual and saccadic execution. Similarly, recovered manual non-decision times are substantially longer (though not more variable) than expected motor execution durations for button presses. These patterns suggest that parts of what the model treats as non-decision time are likely decisional in nature, although perhaps related to "action decision" rather than the "value-based decision" of interest to the authors. To what extent these two processes neatly follow each other or overlap could be usefully considered.

Reviewer #2 (Public review):

Summary:

Zylberberg et al. reanalyze eye-tracking and behavioral data (mostly from Krajbich et al., 2010) to test two predictions of the attentional Drift Diffusion Model, finding that these predictions are not met. Similarly, predictions of normative models (inspired by rational inattention) are not in line with the data, and the authors propose a post-choice model of attention. This model better accounts for the two effects but also does not account for all patterns, so the authors conclude that eye movements most likely reflect both pre- and post-decisional processes.

Strengths:

A clear strength is the systematic falsification-based approach of the paper, establishing (partially) new predictions and testing to what extent these are met by extant models and by a newly developed theory. The authors do a good job in providing intuitions behind the effects and the reasons why models such as the aDDM predict them. The paper is of substantial relevance for the field, as it shows that effects pertaining to the last fixation(s) should be interpreted with caution. Another strength is the paper's transparency as the authors clearly acknowledge that their new model does not do a perfect job either.

Weaknesses:

The paper focuses on analyzing the Krajbich 2010 data, but shows that the second effect replicates in many other datasets. A more principled approach, in which both effects are analyzed and presented for all datasets, would be more convincing. The results should then be shown together for clarity/readability.

Similarly, it would be nice to show to what extent the models' predictions depend (not depend) on using the best-fitting parameter values (are there any parameter settings under which the two effects are not predicted?)

Reviewer #3 (Public review):

Summary:

In this study, the authors reanalyzed choice, RT and gaze datasets collected from human subjects performing a food-choice task. They show that models that posit a causal role for attention in shaping the decision-making process fail to account for empirical observations in the data. These include the attentional drift diffusion model (aDDM) and models that derive attention-choice associations from an optimal policy. The authors show that a model that assumes that gazes are directed towards the chosen option after decision commitment captures more (but not all) empirical findings, suggesting that attention may reflect decisions once they are made instead of contributing to their formation. However, this post-decision-gaze (PDG) model failed to capture all aspects of the data, suggesting that gaze may reflect both decisional and post-decisional operations, and existing models are still missing some features of the gaze-directing process. The authors provide convincing evidence that post-decision gaze explains a number of empirical findings in this task.

Strengths:

(1) The analyses are generally appropriate, and the conclusions are supported by the data.

(2) The study was rigorous, as the authors considered a number of alternative possible models for behavior, and evaluated their performance based on a wide range of qualitative predictions (as opposed to exclusively relying on model comparison).

(3) The proposal that gaze may largely reflect post-decisional processes is interesting, and as far as I am aware, novel.

Weaknesses:

There was limited discussion about why one might allocate attention post-decision. I would have appreciated more discussion on the potential functional consequences or implications of post-decision gaze.

Author response:

The following is the authors’ response to the original reviews.

eLife Assessment

This study combines a five-year field experiment with a meta-analysis to quantify the effects of grazing on ecosystem CO2 fluxes as modified by wetness. The results of this study are potentially valuable, but the Methods description is incomplete and compromises reproducibility and interpretability. A major caveat of this study is that the assessment of CO2 fluxes in time and space is not complete.

We appreciate the comments. The incomplete assessment of CO2 fluxes in time and space has been added in the Limitations and implications for future study Section in the revised manuscript. Method description has been revised as follows:

Lines 205-215, page 7: “Four grazing rotations were conducted each year from 2019 to 2023 in the field study. Before each grazing rotation, two cages (1.2 m × 1.2 m × 1.2 m) were installed in each plot to ensure that the vegetation inside would not be foraged by sheep. Aboveground plant biomass was collected after the end of each grazing rotation. Specifically, in each plot, five 1 m × 1 m quadrats were placed. All aboveground biomass within these quadrats was clipped at ground level and then oven-dried at 65 °C for 48 h to determine the community-level dry biomass. The plant species were classified into C3 and C4 groups (Table S1). Belowground biomass (BGB) was quantified by collecting root biomass using soil core with diameter of 7 cm each September from 2019 to 2023 (Fig. S2). Specifically, two soil cores were collected from each 1 m × 1 m quadrat corresponding to aboveground biomass measurements at depths of 0-30 cm. The belowground parts were then extracted from the soil by washing with water and oven-dried at 65 °C for 48 h to obtain dry weight.”

Lines 243-245, page 8: “The environmental predictors considered in the analysis included grazing intensity, the response ratio of soil temperature and soil moisture, wetness index, and grazing duration.”

Lines 468-470, page 15: “The assessment of ecosystem CO2 fluxes was incomplete due to limited measurement time, area and sampling intervals. This may bias the representation of seasonal cycles and spatial heterogeneity, especially in the regions that was not monitored.”

Public Reviews:

Reviewer #1 (Public review):

Summary:

This study integrates long-term (7th-11th year) ecosystem CO2 flux measurements from a continuously grazed typical steppe in Inner Mongolia with a global meta-analysis of grazing experiments (585 observations) to systematically investigate how the wetness index modulates the effects of grazing intensity on net ecosystem productivity (NEP) in grasslands.

Key findings include:

(1) Heavy grazing significantly reduced gross primary productivity (GPP) and ecosystem respiration (ER) in the typical steppe but did not significantly affect NEP.

(2) Globally, grazing significantly reduced GPP, ER, and NEP, and a higher wetness index and aboveground biomass (AGB) enhance the positive response of NEP to grazing.

(3) Under light and moderate grazing, the response of NEP to grazing was positively correlated with the wetness index, a relationship potentially mediated by a higher plant relative growth rate (RGR) in wetter years.

This study holds considerable practical significance. In the context of global change, comprehending the regulatory function of water is of great importance for the adaptive management of grasslands.

Strengths:

(1) The study cleverly combines long-term in situ observations (revealing temporal dynamics and potential mechanisms) with a global meta-analysis (testing the generality of patterns), forming a complete evidence chain from "process understanding" to "pattern verification".

(2) Focusing on the 7th to 11th years of grazing treatments avoids the common "initial disturbance effects" observed in short-term grazing experiments and truly captures the steady-state response of the ecosystem after it has reached a new equilibrium.

(3) By analyzing plant relative growth rate (RGR) and aboveground biomass (AGB), this study provides mechanistic clues regarding how the wetness index modulates grazing effects. The finding that plant compensatory growth is enhanced in wetter years is a crucial pathway explaining the variation in the NEP response.

(4) The meta-analysis systematically integrates published literature with a substantial sample size (585 observations) and broad geographical coverage (Figure 1B).

(5) The finding that light grazing promotes carbon sinks in wet years, while heavy grazing reduces productivity even in wet years, has direct implications for adaptive grassland management under climate change.

Thank you for the comments. We would like to express our gratitude for your positive and insightful evaluation of this manuscript.

Weaknesses:

Although the paper does have strengths in principle, there are weaknesses and areas for improvement in the paper. In particular:

(1) Incomplete mechanistic chain: While RGR and AGB data offer valuable clues for mechanistic interpretation, the causal chain from "increased wetness → higher RGR → maintained NEP" remains incomplete. Key intermediate processes, such as soil moisture dynamics, nutrient availability, changes in community composition, and leaf photosynthetic physiological parameters, are absent, making the mechanistic explanation somewhat speculative.

Thank you for the comments. The data of the soil moisture dynamics, composition of C3 and C4 plant communities and their species richness have been supplemented to complete the mechanistic chain. However, the structural equation modeling could not be set up if C3 and C4 plant communities were included (Fig. S8). Leaf photosynthetic physiological parameters were not collected in this study. We have made following changes in the revised manuscript:

Lines 339-341, page 11: “In addition, the structural equation model showed that grazing altered net ecosystem productivity by increasing soil temperature and relative growth rate, while decreasing aboveground biomass of plants (Fig. S8).”

Lines 472-474, page 15: “Furthermore, key intermediate processes, such as soil nutrient availability, changes in community composition, and leaf photosynthetic physiological parameters, should also be investigated in the future.”

(2) Inadequate exploration of heterogeneity: The global meta-analysis reveals significant effect heterogeneity. Currently, only a few moderators, such as wetness index, precipitation, and temperature, are analyzed. Potential moderators, including grazing history, grassland type (typical steppe/alpine meadow/savanna), livestock type (cattle/sheep/mixed), and soil type, are not adequately explored.

Thank you for the comments. Grazing duration was presented in Fig. S13. The heterogeneity analysis of grassland types and livestock types was supplemented as Fig. S11 and Fig. S12 in the Supplementary documents. We have also made following changes in the revised manuscript:

Lines 367-370, page 12: “Grazing decreased GPP, ER and NEP in desert and temperate grassland, but had no significant effect on GPP, ER and NEP in alpine grassland and savanna (Fig. S11). In addition, cattle and sheep grazing decreased GPP and NEP, but livestock mixed grazing did not show significant effect on GPP, ER and NEP (Fig. S12).”

(3) Integration of long-term experiment and meta-analysis could be tighter.

Thank you for the comments. Integration of long-term experiment and meta-analysis has been revised in the revised manuscript as follows:

Lines 449-463, page 15: “Our long-term experiment and global meta-analysis jointly provided complementary evidence for understanding the effect of grazing on the ecosystem CO2 fluxes in grassland ecosystems. Long-term grazing field experiments revealed the mechanisms underlying the effects of grazing on ecosystem CO2 fluxes. Global meta-analysis further clarified the universality of these response patterns across different climatic zones and grassland types. It should be noticed that both field experiment and meta-analyses consistently revealed that wetness was an important factor in regulating the ecosystem CO2 fluxes in response to grazing. Under wetter conditions, light grazing promoted compensatory plant growth, enhanced leaf turnover and photosynthetic recovery capabilities, thereby maintaining or even enhancing the NEP (Morgan et al., 2016; Owensby et al., 2006). Conversely, under drier conditions or heavy grazing pressure, a decrease in aboveground biomass and weakened ecosystem resilience jointly limited the ecosystem carbon uptake capacity. These results suggested that wetness modulated the effects of grazing on NEP in grasslands. Moderate grazing may be sustainable under better water conditions, while heavy grazing may exceed the ecosystem resilience threshold even under relatively wet conditions, leading to sustained ecological degradation.”

Reviewer #2 (Public review):

Overgrazing in grasslands is a widespread, enormous problem, and its consequences need more research attention.

However, I have some concerns about the methods in this manuscript.

(1) The plots were grazed from May to September. For the CO2 measurements, it states growing season, and that data was collected on sunny days from 9 to 11 am. However, annual values are reported, and no details on how these were calculated. With no measurements outside of 9 to 11 am, and only on non-sunny days, and only during the growing season, I question how accurate the annual numbers are.

Thank you for the comments. The ecosystem CO2 fluxes were measured during growing season; the description of annual CO2 have been revised to “growing-season” CO2. The ecosystem CO2 fluxes were measured on sunny days during the growing season (May-September) between 9:00 and 11:00 a.m. It can represent typical daytime carbon flux dynamics during the growing season (Fan et al., 2011; Li et al., 2010). In addition, previous studies showed that the ecosystem CO2 fluxes measured at 9:00 and 11:00 a.m. represented the average ecosystem CO2 fluxes of the day (Niu et al., 2008; Rong et al., 2017; Yu et al., 2025), so we use the ecosystem CO2 fluxes measured at 9:00 and 11:00 a.m. to calculate the total ecosystem CO2 fluxes of the day. We have also revised it in the limitation for future study section. We have made following changes in the revised manuscript:

Lines 183-186, page 6: “The gas exchange measurements were conducted on sunny and calm days between 9:00 and 11:00, a time when the ecosystem CO2 represented the daily average (Niu et al., 2008; Rong et al., 2017; Yu et al., 2025), with a frequency of three times per month.”

Lines 468-472, page 15: “The assessment of ecosystem CO2 fluxes was incomplete due to limited measurement time, area and sampling intervals. This may bias the representation of seasonal cycles and spatial heterogeneity, especially in the regions that were not monitored. Multiple time-point sampling should be adopted; the sampling frequency and points should also be increased in the future field study.”

(2) This study uses one chamber, 0.5 by 0.5 m, for a total area of 0.25 m2. This is pretty small, and with no replication, I would like to see more detail on how these are placed, especially when one of the dominant species is a bunchgrass. Sampling on or off a bunchgrass will give very different readings, and neither will be representative of the entire plot. The same for the soil temp and water measurements, there is no detail on how many replicates are in each plot, and working randomly does not work when there are spatial patterns caused by the bunchgrasses in the plots.

Thank you for the comments. Two replicates of the chambers in each plot were set, and their specific sites were labeled in Fig. S2. The study area was located in a typical steppe dominated by Stipa grandis and Leymus chinensis. Therefore, we selected quadrats mainly containing these two species. We have made following changes in the revised manuscript:

Line 186-187, page 6: “In each plot, the chambers of two replicates were measured.”

(3) MAP and MAT are important in this study, and no mention is given if this data is from this site or a neighboring site, and if so, what the distance to this location is.

Thank you for the comments. The data of MAP and MAT were provided in the revised manuscript as follows:

Lines 193-195, page 7: “The precipitation and air temperature were collected from the Xilinhot Meteorological Observation Station, which was near the experimental site in this field experiment.”

(4) Vegetation was sampled in five locations in each plot. I note no detail on the belowground biomass, how many reps, what area, or which depth? This essential data is missing. The same goes for the RGR, and here the authors talk about grazing events. Whereas earlier in section 2.2, it implies continuous grazing every day during the growing season. RGR needs more details, such as how many times, its replication, etc.

Thank you for the comments. The detailed method for how to collect belowground biomass and how to calculate the RGR has been supplemented in the revised manuscript as follows:

Lines 205-215, page 7: “Four grazing rotations were conducted each year from 2019 to 2023 in the field study. Before each grazing rotation, two cages (1.2 m × 1.2 m × 1.2 m) were installed in each plot to ensure that the vegetation inside would not be foraged by sheep. Aboveground plant biomass was collected after the end of each grazing rotation. Specifically, in each plot, five 1 m × 1 m quadrats were placed. All aboveground biomass within these quadrats was clipped at ground level and then oven-dried at 65 °C for 48 h to determine the community-level dry biomass. The plant species were classified into C3 and C4 groups (Table S1). Belowground biomass (BGB) was quantified by collecting root biomass using soil core with diameter of 7 cm each September from 2019 to 2023 (Fig. S2). Specifically, two soil cores were collected from each 1 m × 1 m quadrat corresponding to aboveground biomass measurements at depths of 0-30 cm. The belowground parts were then extracted from the soil by washing with water and oven-dried at 65 °C for 48 h to obtain dry weight.”

(5) Hypothesis 2 is vague: "act as key factors". A more precise hypothesis would be better, otherwise it just leads to p-hacking.

Thank you for the comments. The misleading description has been deleted.

(6) More generally, neither hypothesis is really based on the introduction, as the authors report mixed results in the literature. Hence, the more specific hypothesis 1 reads like it is HARKED, based on the results that the authors found. This is not a good research practice. See the following papers on this topic:

Murphy & Aquinis. 2019. HARKing: How bad can cherry-picking and question trolling produce bias in published results? Journal of Business and Psychology 34:1-17

Bishop. 2019 Rein in the four horses of irreproducibility. Nature 568: 435

Fraser et al. 2018. Questionable research practices in ecology and evolution. Plos One 13:7

Parker et al. 2016. Transparency in ecology and evolution: real problems, real solutions. Trends in ecology and evolution 31:9

Thank you for the suggestion. The hypotheses have been deleted in the revised manuscript to avoid HARKED mistakes.

(7) Figure 2 mentioned n=3, which is correct. However, the variance around the means in the figures is tiny, which raises questions about the replication that is actually used here. I would like to see the entire statistics tables, including DF and sample sizes, included as an appendix, so that the reader can evaluate this much more.

Thank you for the comments. In this study, the data presented in Fig.2 were expressed as mean± standard error (SE):

In addition, we have provided the complete statistical table in Supplementary Table S2 according to the suggestion of the reviewer.

(8) Figure 3, now only low, medium, and high grazing are reported as changes from the control. This can be misleading as the reader can't see how the control varies along the various gradients. I suggest including a figure with the raw data for each as an appendix. In addition, the sample size in 3d, e, f is much higher, and it looks to me like the authors used both the reps and years together. This is not good practice. In addition, full statistics tables should be included in the appendix.

Thank you for the comments. The relationships between ecosystem CO2 fluxes and wetness index have been supplemented in Fig. S6. We have also provided the complete statistical table in Supplementary Table S3.

(9) In Table S1, since there are already a number of recent meta-analyses on this topic, I think there needs to be a stronger justification for this one.

Thank you for the comments. The results of our five-year field monitoring experiments showed that carbon fluxes and their components were significantly influenced by the wetness index. Therefore, we explored whether such effects also occur in grazing experiments at the global scale. The results demonstrated that similar patterns indeed exist worldwide, indicating that our meta-analysis is meaningful and necessary. In addition, previous studies have addressed related topics, net ecosystem productivity has mostly been treated as an auxiliary variable rather than the primary focus of investigation (Jiang et al., 2020; Shi et al., 2022; Zhang et al., 2022; Zhou et al., 2019). In previous studies, meta-analysis literatures on grazing and NEP were limited and not adequate. Our meta-analysis is more comprehensive and reflects the reliability of the results. We have made following changes in the revised manuscript:

Lines 491-492, page 16: “In previous studies, meta-analysis literatures of effects of grazing intensities on NEP were limited and not adequate.”

Lines 512-514, page 16: “In summary, the meta-analysis of this study presents the first comprehensive assessment of how annual wetness index affects the response of ecosystem CO2 fluxes to grazing across global grasslands.”

(10) Regarding the grazing-induced CO2 fluxes, these are only based on the plants and do not incorporate the animal CO2 flux, nor the animal litter CO2 flux, as they were penned at night outside the plots. Thus, the grazing impact is inflated in the data reported here and does not really represent GPP, NET, or RE. This needs to be written about in the discussion section.

Thank you for the comments. The objective of this study was to evaluate the carbon exchange processes from vegetation and soil under grazing disturbance. Thus, the grazing effects reported in this study were the vegetation-soil ecosystem scale rather than the net carbon balance of the entire grazing system. Consequently, our conclusions regarding the effects of grazing on grassland ecosystem carbon exchange remain reliable and ecologically meaningful.

We have made following changes in the manuscript:

Lines 483-490, page 16: “There are also limitations in evaluating the effects of grazing on grassland-livestock ecosystem CO2 fluxes in this study. Specifically, carbon emissions derived from animal respiration were not included in this study. Therefore, the results of the field experiment and meta-analysis in this study should not be interpreted as a complete carbon budget assessment of the grassland-livestock ecosystem. Our primary objective was to evaluate the carbon exchange processes between vegetation and soil under grazing disturbance. Thus, the grazing effects reported in the field experiment and meta-analysis of this study were responses at the vegetation- soil ecosystem scale rather than the net carbon balance of the entire grazing system.”

Reviewer #3 (Public review):

Combining a five-year field experiment with a global meta-analysis, Wu et al. investigate how grazing intensity influences ecosystem carbon dioxide (CO2) fluxes in grasslands and how these effects are regulated by environmental conditions such as grazing duration, wetness index, and soil temperature and moisture responses.

The authors show that the response of net ecosystem productivity (NEP) to light grazing shifts from negative to positive along a wetness gradient, whereas heavy grazing consistently suppresses NEP across wetness conditions. Importantly, this pattern is supported by both the field experiment and the meta-analysis, suggesting that moderate levels of grazing can potentially enhance both plant productivity and carbon sequestration under favorable moisture conditions.

The integration of experimental data with a global synthesis is a particular strength of the study, allowing the authors to evaluate grazing impacts across both temporal variability (precipitation fluctuations in the field experiment) and spatial variability (wetness gradients across global grasslands). Overall, the conclusions are well supported by the data.

However, several aspects of data acquisition, analysis, and presentation could be clarified to further strengthen the reproducibility and interpretation of the results:

(1) A PRISMA-style flow diagram would be helpful for the meta-analysis to clearly illustrate the study selection process and facilitate interpretation of the dataset.

Thank you for the comments. We have added the PRISMA-style flow diagram in the revised supplementary materials. See Fig. S9.

(2) Grazing intensity requires a clearer definition and, where possible, standardization between the field experiment and the studies included in the meta-analysis. Key parameters such as the number of stock per area, days per rotation or per year, and total years of grazing should be clearly defined. In addition, the criteria used to classify grazing intensity into LG, MG, and HG in the meta-analysis should be explicitly described.

Thank you for the comments. In our meta-analysis, the classifications of grazing intensity into light, moderate, and heavy grazing were provided in previous studies. For studies in which grazing intensity was not explicitly reported, we classified grazing intensity according to the USDA criteria (https://www.fs.usda.gov/Internet/FSE_DOCUMENTS/stelprdb5109714.pdf).

The criteria was that: light grazing: approximately equal to a maximum of 40% Utilization (grazing and trampling) of forage standing crop (current and previous years’ growth) at the end of the growing season; moderate grazing: approximately equal to a maximum of 50% Utilization (grazing and trampling) of forage standing crop (current and previous years’ growth) at the end of the growing season; Heavy grazing: greater than 50% Utilization (grazing and trampling) of forage standing crop (current and previous years’ growth) at the end of the growing season (November 15).

We have made following changes in the manuscript:

Lines 271-277, page 9 “(e) The classifications of grazing intensity (light, moderate, and heavy grazing) were primarily based on the definitions provided in the original studies. For studies in which grazing intensity was not explicitly reported, we classified grazing intensity according to the modified Grazing Intensity Classes proposed by the USDA (https://www.fs.usda.gov/Internet/FSE_DOCUMENTS/stelprdb5109714.pdf) (Yin et al., 2023).”

(3) In several sections of the manuscript, it is difficult to distinguish whether phrases such as "this study" or "our study" refer specifically to the field experiment or to the overall study, including both the experiment and the meta-analysis. Clearer wording distinguishing these components would improve readability.

Thanks for the suggestion. The specific distinctions have been revised to clarify whether they refer to the field experiment, meta-analysis or the comprehensive conclusions of the results from both field experiment and meta-analysis.

Lines 50-53, page 2: “Overall, the meta-analysis and field experiment jointly provide global perspectives on the response of ecosystem CO2 fluxes to grazing intensity and improve our knowledge of the factors influencing the response of ecosystem CO2 fluxes to grazing intensity.”

Lines 191-193, page 6: “The annual precipitation and mean annual air temperature (MAT) for the field study area from 2019 to 2023 were obtained from the China Meteorological Data Service Centre (http://data.cma.cn/).”

(4) The discussion attributes the non-significant annual NEP response to intra-annual precipitation variability, with grazing enhancing NEP under wet conditions but suppressing it under dry conditions. Another potential explanation may be that grazing affects GPP and ecosystem respiration (ER) at similar magnitudes (i.e., RR(GPP) ≈ RR(ER)), resulting in limited net changes in NEP.

Thank you for the comments. Another potential explanation has been revised in the manuscript as follows:

Lines 385-389, page 13: “Grazing decreased the responses of ecosystem CO2 fluxes and plant biomass in global grasslands, but only NEP was not significantly affected by grazing in our field experiment (Fig. 4). The lack of a significant response in NEP may be because the site in this field study was managed for year-round continuous low-intensity grazing (Liang et al., 2021). In addition, grazing affected GPP and ER at similar magnitudes, resulting in limited net changes in NEP.”

Recommendations for the authors:

Reviewing Editor Comments:

As you can see from the above reviews and the specific recommendations below, the first issue you should resolve is a full description of methodological detail such that readers can, in principle, repeat your study. They have to know how you placed the chamber to have a representative measure of vegetation (averaging between high and low biomass patches). They also must know how to calculate grazing intensity and assign the values to the three classes. They want to see your raw data and statistical analyses (including formulae, replication, and degrees of freedom). Consider non-linear relationships of grazing and wetness. Try to better integrate the results from the experiment and the meta-analysis. Finally, make sure that you develop hypotheses from the prior knowledge presented in the introduction. For instance, instead of stating that NEP responses would shift from negative to positive with increasing wetness index, simply hypothesize that wetness mitigates the negative effects of grazing on CO2 fluxes, even if you find that this is not true under heavy grazing.

Thank you for the comments. The description of methodological details has been supplemented in the revised manuscript. The results from the experiment and the meta-analysis have been integrated.

Reviewer #1 (Recommendations for the authors):

General suggestions:

(1) The Introduction section and the assumptions should be rewritten and improved. The logicality of the introduction should be revised to better prioritize and contextualize the research problem. The reader is lost since the links between assumptions and previous knowledge are not clear.

Thanks for the suggestion. The Introduction section and the assumptions have been rewritten as follows:

Lines 135-147, page 5: “Previous studies on the effects of grazing intensities on ecosystem CO2 fluxes have been constrained by limited spatial and temporal scales, which has led to an incomplete understanding of how different grazing intensities influence ecosystem CO2 fluxes. Furthermore, does the wetness modulate the effect of grazing on ecosystem CO2 fluxes in the typical steppe? Are these relationships globally generalizable? In this study, we investigated the effects of grazing intensities on ecosystem CO2 fluxes by combining a long-term (7-11 years) field experiment conducted in a typical steppe and a meta-analysis of global grasslands. The objectives of this study were to: (i) investigate the effects of grazing intensity with annual wetness fluctuations on ecosystem CO2 fluxes (GPP, ER and NEP) covering the 7th to 11th years of a continuous grazing experiment in a typical steppe as well as the meta-analysis in global grasslands; (ii) explore how environmental factors (particularly wetness index, soil moisture and temperature, and grazing intensity) regulate the effects of grazing on ecosystem CO2 fluxes.”

(2) In the "Materials and methods" section, you need to provide the reason why you chose the "wetness index" in this study, rather than other drought indices (such as Standardized Precipitation Evapotranspiration Index (SPEI), Aridity Index (AI)).

Thank you for the comments. Although the standardized precipitation evapotranspiration index (SPEI) would be a more appropriate indicator for this study, its calculation requires relatively long and continuous climate data series, which were difficult to obtain in our global meta-analysis. This limitation was particularly important because our study also included a meta-analysis, for which complete climatic datasets were often unavailable from the collected literature. In contrast, the Aridity Index (AI) cannot adequately reflect interannual variability. We have supplemented this limitation in the revised as follows:

Lines 482-483, page 15: “Furthermore, more drought or wetness indices should be investigated in future studies of grazing on ecosystem CO2 fluxes.”

(3) For the results and discussions, the results are currently presented in parallel (long-term experiment first, then meta-analysis). It is recommended to add a dedicated integration paragraph in the discussion.

Thank you for the comments. The dedicated integration paragraph has been supplemented in the revised manuscript as follows:

Lines 450-464, page 15: “Our long-term experiment and global meta-analysis jointly provided complementary evidence for understanding the effect of grazing on the ecosystem CO2 fluxes in grassland ecosystems. Long-term experiments revealed the mechanisms underlying the effects of grazing on plant characteristics and ecosystem CO2 fluxes under control conditions. And global meta-analysis further clarified the universality of these response patterns across different climatic zones and grassland types. It should be noticed that both field experiment and meta-analysis consistently revealed that wetness was an important factor in regulating the effect of ecosystem CO2 fluxes. Under wetter conditions, light grazing promoted compensatory plant growth, enhanced leaf turnover and photosynthetic recovery capabilities, thereby maintaining or even enhancing the NEP. Conversely, under drier conditions or heavy grazing pressure, a decrease in aboveground biomass and weakened ecosystem resilience jointly limited the ecosystem carbon uptake capacity. These results suggested that wetness determined whether grazing promoted or inhibited the carbon sink function in grasslands. Moderate grazing may be sustainable under better water conditions, while heavy grazing may exceed the ecosystem resilience threshold even under relatively wet conditions, leading to sustained ecological degradation.”

(4) Make sure that the whole manuscript has undergone professional proofreading, and check it carefully to avoid language mistakes.

Thank you for the comments. The manuscript has been revised by professional proofreading to avoid language mistakes.

Specific suggestions:

(1) Lines 126-131: It is recommended to explicitly state three levels of research questions: (i) How does grazing intensity affect ecosystem CO2 fluxes? (ii) Does the wetness index modulate this effect? (iii) Are these relationships globally generalizable? This will provide a clear logical thread for the paper.

Thanks for the suggestion. We have made following changes in the revised manuscript:

Lines 135-139, pages 5: “Previous studies on the effects of grazing intensities on ecosystem CO2 fluxes have been constrained by limited spatial and temporal scales, which has led to an incomplete understanding of how different grazing intensities influence ecosystem CO2 fluxes. Furthermore, does the wetness modulate the effect of grazing on ecosystem CO2 fluxes in the typical steppe? Are these relationships globally generalizable?”

(2) Lines 181: Briefly justify the choice of the De Martonne wetness index (Equation 1) in the introduction (why this index over other aridity indices).

Thank you for the comments. The De Martonne wetness index is easier to obtain compared to other drought index, as it only requires annual average temperature and precipitation data for calculation, facilitating the statistics of global meta-analysis. The justification of the De Martonne wetness index has been revised in the manuscript:

Lines 120-125, pages 4-5: “Annual precipitation is one of the climatic parameters, while the wetness index (WI) serves as a more integrative climatic indicator that incorporates both precipitation and temperature, thereby reflecting the overall water surplus or deficit (Song et al., 2019). A higher wetness index (WI > 30) indicates sufficient water availability for plant growth, whereas a lower wetness index (WI ≤ 30) suggests the water availability may be limited (De Martonne, 1926).”

(3) Lines 173-174: Please specify the exact timing of flux measurements (e.g., "measured three times per month between 9:00 and 11:00 AM" is already stated, but add "on sunny and calm days" to ensure consistent conditions).

Thanks for the suggestion. We have supplemented the exact timing of flux measurements in the revised manuscript as follows:

Lines 183-186, page 6: “The gas exchange measurements were conducted on sunny and calm days between 9:00 and 11:00, a time when the ecosystem CO2 fluxes represented the daily average (Niu et al., 2008; Rong et al., 2017; Yu et al., 2025), with a frequency of three times per month.”

(4) The RGR and AGB data provide important clues for explaining the moderating role of the wetness index, but the mechanistic chain can be further refined. It is recommended to use the structural equation model.

Thanks for the comments. The structural equation model has been supplemented to complete the mechanistic chain.

Lines 338-342, page 11: “Heavy grazing decreased C3 plant biomass but increased C4 plant richness compared with other treatments (Fig. S7). In addition, the structural equation model showed that grazing altered net ecosystem productivity by increasing soil temperature and relative growth rate, while decreasing aboveground biomass of plants (Fig. S8).”

Suggested addition: If data on soil moisture, soil nutrients (e.g., ammonium, nitrate), leaf photosynthetic parameters (e.g., maximum photosynthetic rate, stomatal conductance), or community composition are available, please incorporate them into the analysis to test a more complete mechanistic pathway.

Thanks for the comments. The data of the composition of C3 and C4 plant communities and their species richness have been supplemented to complete the mechanistic chain.

Lines 338-342, page 11: “Heavy grazing decreased C3 plant biomass but increased C4 plant richness compared with other treatments (Fig. S7). In addition, the structural equation model showed that grazing altered net ecosystem productivity by increasing soil temperature and relative growth rate, while decreasing aboveground biomass of plants (Fig. S8).”

(5) The current meta - analysis has established the moderating role of the wetness index, but there is room for further exploration: Test for non-linearity. Could the relationship between the wetness index and the grazing effect size (lnRR of NEP) be non-linear in the global data? Consider fitting models that include a quadratic term for the wetness index or using generalized additive models (GAMs). If a threshold is identified, report the threshold estimate and its confidence interval and discuss its management implications.

Thank you for the comments. Following the reviewer's suggestion, we further examined whether there is a nonlinear relationship between the wetness index (WI) and the magnitude of grazing effect (NEP_RR). We compared a linear mixed-effects model including a first-order term for WI with a quadratic mixed-effects model, and set Study ID as a random effect in both models. The results indicated that the AIC value of the quadratic model (256.56) was higher than that of the linear model (241.48), suggesting that adding the second-order term did not improve the model fitting precision. Based on this, we retained the simpler linear model in the revised manuscript.

Note: if the AIC value was lower, the model fitting accuracy was higher.

(6) Section 4.1: This section is quite long. Consider splitting it into 2-3 paragraphs, discussing: (i) overall grazing effects on CO2 fluxes; (ii) the moderating role of the wetness index and its mechanisms; (iii) differential effects of grazing intensities.

Thanks for the suggestion. Section 4.1 has been split into four paragraphs according to the suggestion of the reviewer. (i) overall grazing effects on ecosystem CO 2 fluxes; (ii) the moderating role of the wetness index and its mechanisms; (iii) integrated discussion of field study and meta-analysis.

(7) Integrating the long-term experiment and meta-analysis. Does the effect size observed in the long - term experiment (e.g., an 85.83% increase in NEP under LG in the wettest year) align with the average effect size from the global meta - analysis under similar conditions? If not, what are the potential reasons? (e.g., specificity of the typical steppe, methodological differences between chamber and eddy covariance measurements) Is the mechanism identified in the long - term experiment (e.g., increased RGR) likely to be common globally? What are the joint management implications from both parts of the study? Are there contexts where caution is needed in extrapolating the findings? (e.g., alpine meadows might be more sensitive to grazing).

Thank you for the comments. The long-term experiment and meta-analysis have been integrated in the revised manuscript. The results of field experiment showed that light grazing increased NEP by 85.83%. However, the response of light grazing was not significant in meta-analysis. Both the results of field experiment and global meta-analysis showed that light grazing did not reduce NEP under wetter conditions. The different results in effect size and statistical significance were mainly because the long-term experiment was conducted in a typical grassland ecosystem, which may possess a strong compensatory growth capacity under moderate water conditions. Therefore, light grazing can enhance NEP by increasing plant photosynthetic rate, promoting new leaf growth, and improving community resource utilization efficiency. However, the global meta-analysis integrated different grassland types, climatic conditions and grazing durations. Consequently, the average effect of global meta-analysis may be diluted by the high heterogeneity among ecosystems. The data to calculate RGR were not available in the original studies of the meta-analysis, so we could not include RGR in the meta-analysis. Therefore, it is still unclear if the mechanism of increased RGR could be common globally. The joint management implications have been revised as follows:

Lines 393-400, page 13: “Both the results of field experiment and global meta-analysis showed that light grazing did not reduce NEP under wetter conditions (Fig. 3B, 6B and 7H). Light grazing usually stimulates leaf regrowth following defoliation, and these new leaves often are more physiologically active than the older leaves that contribute much of leaf area in ungrazed treatment (Polley et al., 2008), which likely imply a stronger leaf photosynthesis and C sink (Reich et al., 2007). Considering factors such as different grassland types (Fig. S11), livestock grazing modes (Fig. S12), climatic conditions, and grazing durations, the future grazing studies on NEP should focus more on ANPP and RGR, and extrapolate the results cautiously.”

(8) Figure 8: The conceptual diagram is clear and effectively summarizes the main findings. Briefly explain the meaning of the arrows in the caption.

Thank you for the suggestion. We have revised Fig. 8 accordingly:

“Schematic summary of the effects of grazing intensity on ecosystem carbon dioxide (CO2) fluxes and biomass in global grasslands in the meta-analysis. The blue and black arrows indicate negative effects on grazing and grazing intensities, respectively. Asterisks (*) indicate significant effects on variables at P < 0.05. GPP, gross primary productivity; ER, ecosystem respiration; NEP, net ecosystem productivity; LG, light grazing; MG, moderate grazing; HG, heavy grazing.

(9) Please check all references for consistency with eLife style. Some entries currently have inconsistent formatting (e.g., some include issue numbers, while others do not; page number formatting varies).

Thank you for the comments. We have checked the references one by one to revise them consistent with eLife style.

Reviewer #3 (Recommendations for the authors):

(1) Method citation:

(a) Please cite the original method references rather than studies that applied the methods. For example, the original publication introducing the wetness index (WI) is: De Martonne, E. Une nouvelle fonction climatologique: l'indice d'aridité. La Météorologie 2, 449-458 (1926).

(b) Please also cite the R packages used in the analysis. One straightforward approach is the function citation() in R.

Thank you for the comments. We have cited the references related to the original methods and cited the R packages used in the analysis.

(2) Several expressions would benefit from clarification:

(a) Line 182: What has been "referred to as soil moisture"?

Thank you for the comments. Soil moisture refers to the volumetric water content of the soil, which has been clarified in the revised manuscript as follows:

Lines 199-200, pages 7: “Soil moisture was the soil volumetric water content.”

(b) Line 184-185: Do you mean "soil temperature and soil moisture were measured simultaneously with ecosystem CO2 flux measurements."?

Thank you for the comments. Yes, we simultaneously measured soil temperature and soil moisture using temperature and moisture probes while measuring ecosystem CO 2 flux measurements. We have made following changes in the revised manuscript as follows:

Lines 198-199, page 7: “Soil temperature and moisture at a depth of 0-10 cm were measured simultaneously with ecosystem CO 2 flux measurements, using the probes of the LI-8100 system.”

(c) Line 200: Please clarify what is meant by "the caged plots"?

Thank you for the comments. The misleading words have been deleted in the revised manuscript.

(d) Line 242-244: Do you mean "data from non-grazed treatments were excluded when additional treatments were present"?

Thank you for the comments. Yes, data from non-grazed treatments were excluded when additional treatments were present. We have made following changes in the revised manuscript:

Lines 268-270, page 9: “(c) Data from non-grazed treatments were excluded when additional treatments (e.g., fertilization, experimental warming, or precipitation manipulation) were present.”

(e) Line 263: Should this refer to RR++ instead of RR? Please clarify how RR++ (or lnRR++) was calculated from the reported response ratios and study weights.

Thank you for the comments. RR is the dependent variable of the model, representing the response ratio for each observation. RR++ typically refers to the pooled effect size obtained after all RRs are weighted and subjected to a mixed-effects model, which primarily corresponds to the estimated value of the model intercept β0. We have made following changes in the revised manuscript:

Lines 293-303, page 10: “A linear mixed-effects model, with ‘study’ included as a random factor, was employed to estimate the weighted response ratio (RR++) across studies or within a specific group, fitting with restricted maximum likelihood using the ‘lmer’ function in the ‘lme4’ package (Feng et al., 2023).

where β0 is the coefficient, πstudy denotes the random effect associated with ‘study’ (accounting for autocorrelation among observations from the same study), and ɛ corresponds to the residual sampling error. We checked the normality of the model residuals using the ‘check_normality’ function in the ‘performance’ package. When the assumption of normality was violated, bootstrapping with 999 iterations was performed using the ‘boot’ package to derive the 95% confidence interval (CI) for each RR++ (Chen et al., 2021).”

(3) Line 222: Please list all environmental predictors considered in the analysis.

Thank you for the comments. The environmental predictors considered in the analysis have been listed in the revised manuscript:

Lines: 243-245, page 8: “The environmental predictors considered in the analysis included grazing intensity, the response ratio of soil temperature and soil moisture, wetness index, and grazing duration.”

(4) Line 343: The non-significant response of NEP appears only under MG, while both LG and HG significantly decrease NEP (Fig. 5). It may be helpful to discuss this grazing-intensity-dependent response more explicitly.

Thank you for the comments. The reason why the non-significant response of NEP appeared only under MG was that moderate grazing affected GPP and ER at similar magnitudes, resulting in limited changes in NEP (Fig. 5). However, the reduction in response of GPP was higher than that of ER in LG and HG, resulting in the decrease in NEP. This was consistent with the hypothesis of moderate disturbance, which suggested that ecosystem functions remain stable under moderate levels of disturbance. We have made following changes in the revised manuscript:

Lines 390-393, page 13: “Similarly, the reason why the non-significant response of NEP appeared only under MG may be that moderate grazing affected GPP and ER at similar magnitudes, resulting in limited changes in NEP (Fig. 5). However, the reduction in response of GPP was higher than that of ER in LG and HG, resulting in the decrease in NEP.”

(5) Line 397: The phrase "global-scale study" typically refers to experiments conducted worldwide. "Studies across global grasslands" may be more precise here.

Thank you for the comments. We have made following changes in the revised manuscript:

Lines: 475-477, page 15: “Our meta-analysis has limitations due to the relatively small sample size, which stems from the scarcity of studies across global grasslands exploring the effects of grazing intensity on ecosystem CO2 fluxes.”

(6) Line 403: ...have large amounts of grassland for grazing, "but rarely investigated".

Thank you for the comments. We have made following changes according to the suggestion of the reviewer in the revised manuscript as follows:

Lines 480-482, pages 15-16: “In addition, future studies could conduct more experiments of ecosystem CO2 fluxes in South America, Africa and Oceania, which have large amounts of grasslands for grazing, but were rarely investigated.”

(7) Please remember to cite Figure 8 in the text.

Thank you for the comments. Figure 8 has been cited in the manuscript as follows:

Lines 411-414, page 13: “Moderate grazing decreased GPP and NEP under lower WI, but the responses of GPP and NEP to moderate grazing were similar under higher WI in global grasslands (Figs. 6 and 8), indicating that higher wetness offset the response of GPP and NEP to moderate grazing.”

(8) Figure S1:

(a) The orange line in panel A appears to represent monthly temperature rather than mean annual temperature.

(b) Same for the precipitation, the data shown here should be monthly values rather than annual means.

(c) Consider using a color different from orange for the wetness index in panel C, unless the variable shown is temperature instead.

Thank you for the comments. We have revised Fig. S1 according to the suggestion of the reviewer as follows:

Supplementary page 5: “Fig. S1 The monthly total precipitation and average air temperature (A), annual precipitation (B) and wetness index (C) in the study area of the grazing intensity experiment in the typical steppe from 2019 to 2023.”

(9) Figure S2B: Grazing intensity labels appear in Chinese in the figure. These should be translated into English, and the information on grazing intensity should be provided in the legend or the figure here, as well as in the Method section.

Thank you for the comments. We have replaced the figures included the Chinese labels. See Fig. S2.

(10) Figure S6B: LG significantly affects the relationships between wetness index and ER (P<0.05), but a regression line is missing from the panel.

Thank you for the comments. We have added the regression line between wetness index and ER in Supplementary Fig. S6.

(11) Figure S7A: "Mean" annual precipitation

Thank you for the comments. The “Mean” has been revised in Figure S10A.

References

Fan, Y., Zhang, X., Wang, J., & Shi, P. (2011). Effect of solar radiation on net ecosystem CO2 exchange of alpine meadow on the Tibetan Plateau.Journal of Geographical Sciences, 21(4), 666-676.

Jiang, Z., Hu, Z., Lai, D., Han, D., Wang, M., Liu, M., Zhang, M., & Guo, M. (2020). Light grazing facilitates carbon accumulation in subsoil in Chinese grasslands: A meta-analysis. Global Change Biology, 26(12), 7186–7197.

Li, X., Fu, H., Guo, D., Li, X., & Wan, C. (2010). Partitioning soil respiration and assessing the carbon balance in a Setaria italica (L.) Beauv. Cropland on the Loess Plateau, Northern China. Soil Biology and Biochemistry, 42(2), 337-346.

Niu, S., Wu, M., Han, Y., Xia, J., Li, L., & Wan, S. (2008). Water‐mediated responses of ecosystem carbon fluxes to climatic change in a temperate steppe. New Phytologist, 177(1), 209-219.

Rong, Y., Johnson, D. A., Wang, Z., & Zhu, L. (2017). Grazing effects on ecosystem CO2 fluxes regulated by interannual climate fluctuation in a temperate grassland steppe in northern China. Agriculture, Ecosystems & Environment, 237, 194-202.

Shi, R., Su, P., Zhou, Z., Yang, J., & Ding, X. (2022). Comparison of eddy covariance and automatic chamber‐based methods for measuring carbon flux. Agronomy Journal, 114.

Wan, L., Liu, G., & Su, X. (2025). Global meta-analysis reveals different grazing management strategies change greenhouse gas emissions and global warming potential in grasslands. Geography and Sustainability, 6(3), 100251.

Yin, M., Gao, X., Kuang, W., & Tenuta, M. (2023). Soil N2O emissions and functional genes in response to grazing grassland with livestock: A meta-analysis. Geoderma, 436, 116538.

Yu, H., Wang, X., Wu, Y., Wang, C., Yan, R., Xu, D., & Xin, X. (2025). Light grazing tends to enhance ecosystem carbon sequestration and resource use efficiency in a meadow steppe of northern China. Agricultural and Forest Meteorology, 372, 110690.

Zhang, R., Tian, D., Chen, H. Y. H., Seabloom, E. W., Han, G., Wang, S., Yu, G., Li, Z., & Niu, S. (2022). Biodiversity alleviates the decrease of grassland multifunctionality under grazing disturbance: A global meta-analysis. Global Ecology and Biogeography, 31(1), 155–167.

Zhou, G., Luo, Q., Chen, Y., Hu, J., He, M., Gao, J., Zhou, L., Liu, H., & Zhou, X. (2019). Interactive effects of grazing and global change factors on soil and ecosystem respiration in grassland ecosystems: A global synthesis. Journal of Applied Ecology, 56(8), 2007–2019.

Version 2

Abstract

Overgrazing has induced widespread grassland degradation globally. However, how grazing intensity affects ecosystem carbon dioxide (CO2) fluxes with wetness fluctuations in grasslands at a global scale remains poorly understood. Here we measured ecosystem CO2 fluxes covering the 7th to 11th years of a continuous grazing experiment in a typical steppe, and conducted a meta-analysis of 585 observations by collecting relevant data in global grasslands. Our experimental results showed that mean gross primary productivity (GPP) and ecosystem respiration (ER) were decreased under heavy grazing (HG) compared with control without grazing (CK), whereas net ecosystem productivity (NEP) was not different between HG and CK in the typical steppe. The NEP was higher under light grazing in the wetter years, but it was not different between LG and CK under drier years in the typical steppe. The structural equation model showed that grazing altered NEP by increasing soil temperature and relative growth rate, while decreasing AGB. In contrast, grazing decreased GPP, ER and NEP, respectively, in global grasslands. Both our field experiment and the meta-analysis revealed that the response of NEP to light grazing, rather than heavy grazing, was lineally correlated with the wetness index. Higher wetness and aboveground biomass (AGB) increased the response of NEP to grazing in global grasslands. However, heavy grazing reduced NEP and AGB even under higher wetness indices in global grasslands, resulting from a loss of their resilience in long-term heavy grazing. These findings indicate that light grazing appears to be a promising management to promote plant relative growth rate and carbon sequestration. Overall, the meta-analysis and field experiment jointly provide global perspectives on the response of ecosystem CO2 fluxes to grazing intensity and improve our understanding of the factors influencing the response of ecosystem CO2 fluxes to grazing intensity.

Introduction

Globally, grassland ecosystems cover approximately one fifth of the world land surface and contain about 20-30% of the global pool of soil organic carbon (SOC) (Du et al., 2022). Grasslands also play a vital role in livestock production, promoting carbon (C) sequestration and regulating climate change (Heimann & Reichstein, 2008; Zhou et al., 2017). Grazing is a widespread land-use activity that alters terrestrial C dynamics (He et al., 2020; Zhou et al., 2025). However, grassland ecosystems have seriously deteriorated globally (Zhang et al., 2025). Grassland degradation is influenced by a multitude of driving factors, such as overgrazing and precipitation limitation (Sloat et al., 2018). Furthermore, the C balance of grasslands is also increasingly influenced by changes in grazing management practices and water availability (Hilker et al., 2014; Lemoine et al., 2016; Ren et al., 2025). Overgrazing and drier conditions could potentially drive grasslands from acting as C sinks to C sources (Hao et al., 2018; Morgan et al., 2016; Shao et al., 2013).

Gross primary productivity (GPP) and ecosystem respiration (ER) are critical processes governing C sequestration and emission in terrestrial ecosystems (Chen et al., 2023; Knauer et al., 2023). Net ecosystem productivity (NEP) is the balance between ecosystem C uptake and release, and serves as a key indicator of the net C exchange (Niu et al., 2010; Zhang et al., 2024). Grazing influences NEP by altering GPP and ER (Fig. 1). Grazing directly removes aboveground plant tissues through herbivory, altering vegetation cover, soil temperature, and aboveground biomass (AGB) and thereby influencing plant photosynthesis and GPP (Owensby et al., 2006; Zhou et al., 2007). Plants with high relative growth capacity may rapidly restore leaf area and photosynthesis tissues after grazing, potentially maintaining or even enhancing GPP (Oesterheld & McNaughton, 1991; Penner & Frank, 2021). In contrast, heavy grazing or water limitation may constrain plant regrowth, limit soil moisture, leaf area recovery, and reduce GPP. Grazing may also affect ER by reducing litter and root carbon inputs, altering belowground carbon allocation, and modifying soil bulk density, porosity, water infiltration, and aeration through livestock tamping, with consequences for root respiration and microbial decomposition (Bardgett & Wardle, 2003; Wu et al., 2023; Xun et al., 2018). Variation in NEP responses to grazing may depend on whether post-grazing recovery of GPP can compensate for carbon losses through ER. Thus, plant relative growth rate and compensatory growth capacity may represent important mechanistic traits linking grazing disturbance, photosynthesis recovery, GPP dynamics, and NEP response.

Conceptual illustration of how grazing with wetness fluctuation would affect net ecosystem productivity (A), and geographic (B) and climatic distributions of data (C) included in this meta-analysis.

Assessing ecosystem carbon dioxide (CO2) fluxes (GPP, ER and NEP) in grasslands is essential to the terrestrial C cycle at local, regional and global scales. However, current evidence on the effect of grazing on ecosystem CO2 fluxes in grasslands remains inconsistent, with studies reporting both positive (Chang et al., 2020; Du et al., 2022) and negative (Liu et al., 2020; Wang et al., 2023; Zhu et al., 2015) effects. For instance, some studies demonstrated that grazing decreased GPP and ER (Nakano et al., 2020; Qi et al., 2024; Wang et al., 2025), while others reported an increase or no difference in GPP or ER (Chang et al., 2020; Luo et al., 2015; Shi et al., 2022). Similarly, some studies investigated that grazing significantly reduced NEP (Wang et al., 2023; Zhu et al., 2015), while others reported an increase or no difference in NEP under grazing (Du et al., 2022; Gomez-Casanovas et al., 2018; Yu et al., 2025). The variations in the responses of ecosystem CO2 fluxes are likely driven by differences in grazing intensities, climatic variations, livestock and grassland types (Hou et al., 2016; Wang et al., 2023; Xiao et al., 2011; Zhu et al., 2015).

The effects of grazing intensities on ecosystem CO2 fluxes in global grasslands were also varied in different studies (Jiang et al., 2020; Rong et al., 2017; Zhang et al., 2022). Specifically, heavy grazing usually reduced the GPP, ER and NEP (Ju et al., 2024; Rong et al., 2017; Zhang et al., 2022), but a meta-analysis also reported the response of NEP was increased in Chinese grasslands (Jiang et al., 2020). Generally, light grazing can enhance NEP by promoting plant regeneration and improving photosynthetic efficiency, whereas moderate to heavy grazing tends to reduce NEP, largely as a result of decreased vegetation coverage and plant biomass (Yu et al., 2025). Conversely, a meta-analysis showed that the response of NEP was decreased under light and moderate grazing (Zhang et al., 2022). However, it was reported that moderate grazing led to the highest NEP (Zhu et al., 2015). These conflicting results underscore the complex responses of ecosystem CO2 fluxes in grasslands to grazing and highlight the necessity to identify key factors controlling ecosystem CO2 flux dynamics under grazing management. This understanding is essential for accurately modelling and predicting grassland C dynamics under global change.

Many studies have demonstrated that interannual variation in grassland NEP can be substantial, primarily driven by fluctuating climatic conditions, as they can directly regulate vegetation growth via altering soil moisture and temperature (Polley et al., 2008; Rong et al., 2017; Shen et al., 2016). Annual precipitation is one of the climatic parameters, while the wetness index (WI) serves as a more integrative climatic indicator that incorporates both precipitation and temperature, thereby reflecting the overall water surplus or deficit (Song et al., 2019). A higher wetness index (WI > 30) indicates sufficient water availability for plant growth, whereas a lower wetness index (WI ≤ 30) suggests the water availability may be limited (De Martonne, 1926). Considerable uncertainty remains regarding how NEP responds to grazing under different water conditions. Although evidence suggested that, under wet year, grazing significantly increased NEP; under dry year, grazing suppressed NEP and can even turn NEP negative (Shao et al., 2013). In general, grazing tends to have a less suppressive effect on NEP under wet than dry conditions (Okach et al., 2019; Zhou et al., 2019). For example, heavy grazing consistently reduced NEP, with a more pronounced decline occurring during dry years in a desert steppe (Jin et al., 2023). In addition, overgrazing led to decoupling of precipitation patterns and ecosystem carbon exchange in the desert steppe, primarily by altering plant community composition (Wang et al., 2023). However, it remains unclear how NEP responds to grazing intensities under water availability fluctuations across global grasslands.

Previous studies on the effects of grazing intensities on ecosystem CO2 fluxes have been constrained by limited spatial and temporal scales, which has led to an incomplete understanding of how different grazing intensities influence ecosystem CO2 fluxes. Furthermore, does the wetness modulate the effect of grazing on ecosystem CO2 fluxes in the typical steppe? Are these relationships globally generalizable? In this study, we investigated the effects of grazing intensities on ecosystem CO2 fluxes by combining a long-term (7-11 years) field experiment conducted in a typical steppe and a meta-analysis of global grasslands. The objectives of this study were to: (i) investigate the effects of grazing intensity with annual wetness fluctuations on ecosystem CO2 fluxes (GPP, ER and NEP) covering the 7th to 11th years of a continuous grazing experiment in a typical steppe as well as the meta-analysis in global grasslands; (ii) explore how environmental factors (particularly wetness index, soil moisture and temperature, and grazing intensity) regulate the effects of grazing on ecosystem CO2 fluxes.

Materials and methods

Experimental site

The grazing experiment was conducted at Xilinhot National Climate Observatory (44°12′ N, 116°19′ E, 1129 m above sea level), in the middle part of Inner Mongolian typical steppe in China. The average annual precipitation was 291 mm from 2019 to 2023, with average precipitation in the growing season (May to September) varying from 168 to 325 mm (Fig. S1). The average mean air temperature was 3.9 (from 2019 to 2023). The soil is classified as a Haplic Calcisol based on the FAO soil classification system or a Calcic-Orthic Aridisol based on the USDA soil classification system (Liang et al., 2021). The vegetation community is dominated by Stipa grandis P. Smirn (perennial bunchgrass) and Leymus chinensis Trin. Tzvel (perennial rhizomatous grass) (Wu et al., 2024).

Experimental design

The grazing experiment was initiated in 2013 and was based on a randomized design (Fig. S2). There were four treatments with three replicates: control without grazing (CK), light grazing (LG, 0.64 sheep ha-1 year-1), moderate grazing (MG, 1.28 sheep ha-1 year-1) and heavy grazing (HG, 2.56 sheep ha-1 year-1) conducted in plots with 1.44 ha (120 × 120 m2) established in the research station (Fig. S2). Two-year-old sheep similar in size were selected to use for grazing from May to September in each year. Sheep were grazing from 7:00 AM to 6:00 PM each day, and then gathered into an enclosure at night (Liang et al., 2021).

Measurement of ecosystem CO2 fluxes

Ecosystem CO2 fluxes were measured using a transparent chamber (0.5 × 0.5 × 0.8 m3) connected to an infrared gas analyzer (LI-8100, Lincoln, NE, USA). To ensure an airtight seal during measurements, the junction between the chamber bottom and the aluminum base was secured with tapes. After positioning the chamber, the enclosed air was continuously mixed by two fans in the chamber. The CO2 concentrations were recorded 12 times at 10-second intervals over a 2-minute period. The NEP was then determined from the rate of CO2 concentration change, combined with chamber volume and headspace temperature data. A positive NEP value indicated net ecosystem CO2 uptake, whereas a negative value denoted net ecosystem CO2 release. Following each NEP measurement, the chamber was removed to vent the headspace and then covered with a dark cloth to inhibit photosynthesis before being repositioned on the same base for ER measurement. Both the NEP and ER were measured throughout the growing seasons from 2019 to 2023, corresponding to the 7th to 11th years of grazing. The GPP was thus calculated as the sum of NEP and ER using data collected over this five-year period. The gas exchange measurements were conducted on sunny and calm days between 9:00 and 11:00, a time when the ecosystem CO2 fluxes represented the daily average (Niu et al., 2008; Rong et al., 2017; Yu et al., 2025), with a frequency of three times per month. In each plot, the chambers of two replicates were measured. Thus, six replicates per treatment were set up in the field experiment.

Climate conditions and soil microclimate

The annual precipitation and mean annual air temperature (MAT) for the field study area from 2019 to 2023 were obtained from the China Meteorological Data Service Center (http://data.cma.cn/). The precipitation and air temperature were collected from the Xilinhot Meteorological Observation Station, which was near the experimental site in this field experiment. The De Martonne wetness index (WI) was calculated as follows (De Martonne, 1926; Song et al., 2019): (1) Wetnessindex(WI)=MAPMAT+10 Soil temperature and moisture at a depth of 0-10 cm were measured simultaneously with ecosystem CO2 flux measurements, using the probes of the LI-8100 system. Soil moisture was the soil volumetric water content. Measurement sites for soil temperature and moisture were selected near the collars randomly placed in the plots. They were investigated synchronized with the measurements of ecosystem CO2 fluxes.

Plant sampling and analyses

Four grazing rotations were conducted each year from 2019 to 2023 in the field study. Before each grazing rotation, two cages (1.2 m × 1.2 m × 1.2 m) were installed in each plot to ensure that the vegetation inside would not be foraged by sheep. Aboveground plant biomass was collected after the end of each grazing rotation. Specifically, in each plot, five 1 m × 1 m quadrats were placed. All AGB within these quadrats was clipped at ground level and then oven-dried at 65 °C for 48 h to determine the community-level dry biomass. The plant species were classified into C3 and C4 groups. Belowground biomass (BGB) was quantified by collecting root biomass using soil core with diameter of 7 cm each September from 2019 to 2023 (Fig. S2). Specially, two soil cores were collected from each 1 m × 1 m quadrat corresponding to BGB measurements at depths of 0-30 cm. The belowground parts were then extracted from the soil by washing with water and oven-dried at 65 °C for 48 h to obtain dry weight.

The relative growth rate (RGR) was defined as the ratio of aboveground biomass regeneration to the number of days between two consecutive grazing events and was calculated by Eq. (2) (Zong & Shi, 2019): (2) RGR=(AGBEi+1−AGBi)Δt where AGBEi+1: the AGB inside the cages after the (i+1)-th grazing round, AGBi: the remaining AGB after the i-th grazing, △t: the number of days between the i-th and (i+1)-th grazing rounds.

Aboveground net primary productivity (ANPP) was estimated as the sum of the total herbage consumed by livestock throughout the growing season and the AGB measured at the end of the final grazing rotation of the season.

Data Analysis

All the analyses were conducted using the R (version 4.4.3) (Team, 2023). Annual mean values were calculated for GPP, ER, NEP, AGB, BGB, root/ shoot ratio, soil temperature, and soil moisture from 2019 to 2023. Differences across years and among treatments within each year were examined using the Least Significant Difference (LSD) test. Response ratios of GPP, ER, NEP, AGB, BGB, root/ shoot ratio, soil temperature, and soil moisture were calculated based on Eq. (3). Additionally, linear regression analysis was conducted using the “lm” function to evaluate the relationships of GPP, NEP, AGB, BGB, and RGR with WI. (3) RR=ln⁡(XTXC)=ln⁡(XT)−ln⁡(XC) Where XT and XC are the mean values of the abovementioned indicators for the grazing (T) and control (C) treatments, respectively.

The repeated measures ANOVA (the ezANOVA function of the “ez” R package) was used to examine the effects of grazing treatment and sampling years on ecosystem CO2 fluxes (GPP, ER, and NEP), microclimatic variables (soil temperature and moisture), and plant biomass parameters (AGB, BGB, and root/ shoot ratio) across the five-year study period from 2019 to 2023 (Chen et al., 2024).

We also conducted a random forest model to identify the significant environmental predictors of GPP, ER, NEP, AGB and BGB. The environmental predictors considered in the field study included grazing intensity, the response ratio of soil temperature and soil moisture, wetness index, and grazing duration. The random forest model was implemented in R (version 4.4.3) using the “randomForest” and “rfPermute” packages. In this model, the importance of each predictor was evaluated based on the percentage increase in the mean squared error (MSE) when that predictor was permuted; a higher MSE value indicated greater importance of the factor. The MSE values for each decision tree, based on out-of-bag estimates from the random forest model, were generated using the “rfPermute” package to assess the relative importance of the predictor variables (Liao et al., 2024). To evaluate the direct and indirect effects of grazing intensity, wetness index, soil temperature, soil moisture, RGR, AGB, and C3and C4 plant biomass and richness on NEP, we conducted a structural equation modeling (SEM) analysis using the ‘piecewiseSEM’ packages. A good fit of the SEM was evaluated using Fisher’s C value, which was non-significant (P > 0.05) (Yang et al., 2025).

Meta-analysis of grazing experiments in global grasslands

To assess the effects of grazing on GPP, ER, and NEP in global grasslands, we conducted a systematic literature search for relevant papers published up to June 8, 2025, using the Web of Science (http://apps.webofknowledge.com/) and China National Knowledge Infrastructure (CNKI, https://www.cnki.net). The keywords and terms used were (“grazing” OR “stocking” OR “livestock” OR “grazing intensity”) AND (“net ecosystem productivity” OR “net ecosystem exchange” OR “gross ecosystem productivity” OR “ecosystem respiration” OR “soil carbon flux” OR “soil CO2” OR “soil carbon flux” or “soil carbon emission”). Relevant studies were selected based on the following criteria (Fig. S9): (a) Studies were conducted using simulating grazing experiments, and studies with trampling and mowing were excluded. (b) The control and grazing plots were under similar conditions, such as dominant species, community composition, soil type and climate. (c) Data from non-grazed treatments were excluded when additional treatments (e.g., fertilization, experimental warming, or precipitation manipulation) were present. (d) The mean and sample size for any measurements of GPP, ER, and NEP under both grazing and control conditions must be reported. (e) The classifications of grazing intensity (light, moderate, and heavy grazing) were primarily based on the definitions provided in the original studies. For studies in which grazing intensity was not explicitly reported, we classified grazing intensity according to the modified Grazing proposed by the USDA (https://www.fs.usda.gov/Internet/FSE_DOCUMENTS/stelprdb5109714.pdf) (Yin et al., 2023). Finally, 62 published papers with 585 observations of GPP, ER, and NEP responses (n = 161, 249, and 175, respectively) were selected (Fig. 4), and the geographical location of these observations were presented in Fig. 1. Meanwhile, the study site, elevation, MAP, MAT, and grazing intensity from the selected papers were recorded. The wetness indices based on MAP and MAT of each site were calculated according to Eq. (1). Where available, data for soil temperature, soil moisture, AGB, BGB, and root/ shoot ratio were also recorded.

In addition, to better explore the effects of grazing intensity and environmental factors on ecosystem CO2 fluxes, the data was classified into different subgroups by grazing intensity (LG, MG, and HG), mean annual temperature (≤5 and >5), wetness index (≤ 30 and > 30), and precipitation (≤ 400 mm and > 400 mm), grassland type (desert grassland, temperate grassland, alpine grassland, savanna, and others), and livestock type (cattle, sheep, and mix).

The effect size of grazing on each variable was quantified using the natural logarithm of response ratios (RR) based on Eq. (3) (Feng et al., 2023).

The weighing of the ith study (WRR) was calculated as (Pittelkow et al., 2015): (4) WRR=(NC×NT)(NC+NT) where NT and NC are the replicate numbers of treatment and control, respectively.

A linear mixed-effects model, with ‘study’ included as a random factor, was employed to estimate the weighted response ratio (RR++) across studies or within a specific group, fitting with restricted maximum likelihood using the ‘lmer’ function in the ‘lme4’ package (Feng et al., 2023). (5) RR=β0+πstudy+ε where β0 is the coefficient, πStudy denotes the random effect associated with ‘study’ (accounting for autocorrelation among observations from the same study), and corresponds to the residual sampling error. We checked the normality of the model residuals using the ‘check_normality’ function in the ‘performance’ package. When the assumption of normality was violated, bootstrapping with 999 iterations was performed using the ‘boot’ package to derive the 95% confidence interval (CI) for each RR++ (Chen et al., 2021). The percentage change (%) of the response ratio was calculated as follows: (6) PercentageChange=(eRR+++−1)×100% The lnRR++ was considered statistically significant if its 95% confidence interval (CI) did not overlap zero (Chen et al., 2021).

The random forest model was conducted using “randomForest” and “rfPermute” packages to identify the relative importance of the predictor variables (wetness index, grazing duration and intensity, the responses of soil moisture and temperature and livestock types) of GPP, ER, NEP, AGB and BGB in the meta-analysis (Liao et al., 2024).

Results

Grazing intensity effects on ecosystem CO2 fluxes and plant biomass in the field experiment

Continuous heavy grazing reduced the mean GPP by 29.87% and the mean ER by 30.94% in the grazing experiment; however, NEP did not differ between HG and CK (Fig. 2). The mean GPP and ER were decreased by 17.46% and 22.72%, respectively, in MG than in CK, but NEP was not different between MG and CK (Table S2). The mean ER was decreased by 21.49% and 29.84%, respectively, in MG and HG than in LG. The mean NEP was decreased by 37.41% in HG than in LG, but was not different among LG, MG and CK. The GPP and NEP were 29.27% and 85.83% higher, respectively, in LG than in CK, but the ER was not different in 2020, the year with the highest precipitation and wetness index among the five years (Fig. S1 and 2). The mean aboveground and belowground biomass was decreased by 39.27% and 19.44%, respectively, in HG than in CK, but the mean root shoot ratio was increased by 30.59% in HG than in CK (Fig. 2; P < 0.05).

Effects of different grazing intensities on ecosystem carbon dioxide (CO2) fluxes and biomass in our field experiment.

CK, enclosure; LG, light grazing; MG, moderate grazing; HG, heavy grazing. Data (means ± SE, n = 3) followed by different lowercase letters indicate differences at P < 0.05 between treatments.

Wetness was the most important predictor of NEP, while grazing intensity was the most important predictor of GPP and ER, revealing a strong coupling between CO2-driven changes in wetness index and grazing intensity (Fig. S3). Grazing intensity and duration were also the important predictors influencing AGB and BGB. The linear regression analysis showed that overall grazing as well as light grazing stimulated the responses of GPP and NEP with increasing wetness index (Fig. 3 and Table S3; P < 0.05). This relationship could be explained by the significant positive correlation between NEP and both ANPP and RGR (Fig. S4). In the wettest year of 2020, the relative growth rate peaked relative to other years, reaching levels 3.16, 2.76 and 1.96 times higher under LG, MG, and HG respectively, compared to CK (Fig. S5). The NEP increased with WI under HG (Fig. S6; P < 0.01), whereas the response ratio of NEP to heavy grazing was not significantly correlated with WI (Fig. 3B). Grazing reduced the response of GPP, ER, AGB, BGB and soil moisture in the field experiment (Fig. 4A). Heavy grazing decreased C3 plant biomass but increased C4 plant richness compared with other treatments (Fig. S7). In addition, the structural equation model showed that grazing altered net ecosystem productivity by increasing soil temperature and relative growth rate, while decreasing aboveground biomass of plants (Fig. S8).

Relationships of the responses of gross primary productivity (GPP) and net ecosystem productivity (NEP) with potential regulating factors from 2019 to 2023 under different grazing intensity.

Overall, overall grazing included light, moderate and heavy grazing; LG, light grazing; MG, moderate grazing; HG, heavy grazing. The significant regression lines and 95% confidence intervals are shown with lines (solid for significant) and shaded areas, respectively.

Synthesis of grazing on ecosystem CO2 fluxes and plant biomass across global grasslands

In the meta-analysis across global grasslands, grazing significantly decreased ecosystem CO2 fluxes by 13.56% to 30.46% (Fig. 4B). Heavy grazing had a more pronounced negative effect on ecosystem CO2 fluxes compared to light and moderate grazing (Fig. 5A). The response of GPP was decreased by 15.89%, 11.88% and 26.34%, respectively, under LG, MG and HG. The response of ER was decreased by 7.42%, 14.91% and 21.82%, respectively, under LG, MG and HG. The response of NEP was decreased by 18.86% and 34.89%, respectively, in LG and HG, but was not significant in MG. Light, moderate and heavy grazing reduced AGB by 27.89%, 34.57% and 66.17% (Fig. 5B). Moderate and heavy grazing decreased the BGB by 14.85% and 26.13%, respectively, but the response of BGB was not significant under light grazing. Grazing increased the root shoot ratio by 31.78% to 111.11%. Heavy grazing had a more pronounced positive effect on root/ shoot ratio compared to light and moderate grazing.

Effects of grazing on gross primary productivity, ecosystem respiration, net ecosystem productivity, aboveground biomass, belowground biomass and root shoot ratio, soil moisture and soil temperature, in global grasslands in the field experiment in this study (A) and from the meta-analysis (B).

Circles and error bars represent average parameter estimates and 95% confidence interval (CI). The star (*) indicates significance when the CI did not overlap zero. Sample sizes of observations for each variable are displayed on the left.

Effects of different grazing intensities on ecosystem CO2 fluxes, biomass and soil temperature and moisture in global grasslands from the meta-analysis.

Circles and error bars represent average effects and their 95% confidence interval (CI). The star (*) indicates significant effects, i.e., the 95% CI did not overlap zero. The sample size for each variable is shown on the left. LG, light grazing; MG, moderate grazing; HG, heavy grazing.

Globally, grazing decreased GPP, ER and NEP under lower precipitation conditions (≤ 400 mm), while the reductions were not significant under higher precipitation conditions (> 400 mm) (Fig. S10A). Grazing significantly reduced GPP, ER, and NEP at mean annual temperatures ≤ 5, but these effects were not significant at temperatures > 5 (Fig. S10B). Similarly, under lower wetness conditions (WI ≤ 30), grazing decreased GPP, ER, and NEP more than under higher wetness conditions (WI > 30) (Fig. 6). The response of NEP was decreased under lower WI (WI ≤ 30) in MG and HG, but it was not different under higher WI (WI > 30) in MG. The response of NEP was decreased under higher WI (WI > 30) in HG. Higher wetness index mitigated the response of NEP to grazing. The duration of grazing also influenced its effect: GPP and ER declined more substantially under longer grazing durations (Fig. S10C). The most pronounced decline in NEP occurred under grazing durations of 5 to 10 years (P < 0.001). Grazing decreased GPP, ER and NEP in desert and temperate grassland, but had no significant effect on GPP, ER and NEP in alpine grassland and savanna (Fig. S11). In addition, cattle and sheep grazing decreased GPP and NEP, but livestock mixed grazing did not show significant effect on GPP, ER and NEP (Fig. S12).

Responses of ecosystem CO2 fluxes and biomass to grazing (A) and different grazing intensities (B-C), at wetness index ≤ 30 and wetness index > 30 in global grasslands from the meta-analysis.

Circles and error bars represent average parameter estimates and 95% confidence interval (CI). The sample size for each variable is shown on the left. LG, light grazing; MG, moderate grazing; HG, heavy grazing.

In global grasslands, the key environmental factors influencing GPP and ER response to grazing included WI, grazing duration, the response of soil moisture, and grazing intensity (Fig. S13; P < 0.05). For NEP, the major predictors were WI, grazing duration and intensity (P < 0.05). The most important environmental parameters for AGB response to grazing were grazing intensity, WI, grazing duration, the response of soil moisture and temperature and livestock types. For BGB, the major predictors were WI, grazing duration, the response of soil moisture and temperature. The response of NEP was positively correlated with WI under light, moderate and total grazing in global grasslands (Fig. 7; P<0.1). The response of ecosystem CO2 fluxes was affected by AGB under grazing in global grasslands (Fig. S14).

Relationships of the responses of ecosystem CO2 fluxes with potential regulating factors in global grasslands from the meta-analysis.

The marginal (m) and conditional (c) R2 indicate the proportion of variance explained by the fixed effects and by both the fixed and random effects (study), respectively. The black solid lines and gray shaded areas in panels are the mean and 95% confidence interval of the slope, respectively. The sizes of points are proportional to their corresponding weights. RR, response ratio; overall, total grazing; LG, light grazing; MG, moderate grazing; HG, heavy grazing.

Discussion

The responses of ecosystem CO2 fluxes and plant biomass to grazing and controlling factors

Grazing decreased the responses of ecosystem CO2 fluxes and plant biomass in global grasslands, but only NEP was not significantly affected by grazing in our field experiment (Fig. 4). The lack of a significant response in NEP may be because the site in this field study was managed for year-round continuous low-intensity grazing (Liang et al., 2021). In addition, grazing affected GPP and ER at similar magnitudes, resulting in limited net changes in NEP. Similarly, the reason why the non-significant response of NEP appeared only under MG may be that moderate grazing affected GPP and ER at similar magnitudes (Fig. 5). However, the reduction in response of GPP was higher than that of ER in LG and HG, resulting in the decrease in NEP. Both the results of field experiment and global meta-analysis showed that light grazing did not reduce NEP under wetter conditions (Fig. 3B, 6B and 7H). Light grazing usually stimulates leaf regrowth following defoliation, and these new leaves often are more physiologically active than the older leaves that contribute much of leaf area in ungrazed treatment (Polley et al., 2008), which likely imply a stronger leaf photosynthesis and C sink (Reich et al., 2007). Considering factors such as different grassland types (Fig. S11), livestock grazing modes (Fig. S12), climatic conditions, and grazing durations, the future grazing studies on NEP should focus more on ANPP and RGR, and extrapolate the results cautiously.

Grazing stimulated the NEP under wet conditions, but depressed it under dry conditions, a pattern consistent across the typical steppe in our experiment and supported by our global meta-analysis (Figs. 3B and 7H). This finding aligned with the finding that the response of NEP increased with the increase in wetness index under grazing. This relationship could be explained by the significant positive correlation of NEP with AGB (Fig. S14). The linear regression analysis also showed that light grazing stimulated the responses of NEP with increasing wetness index. This relationship could be explained by the significant positive correlation between NEP and both ANPP and RGR (Fig. S4). Under lower wetness conditions (WI ≤ 30), light grazing reduced GPP and ER to a lesser extent than moderate and heavy grazing (Fig. 6B). Compared with higher grazing intensities, light grazing consumed less forage by livestock, thereby better mitigated the effect of lower moisture on GPP and ER. Moderate grazing decreased GPP and NEP under lower WI, but the responses of GPP and NEP to moderate grazing were similar under higher WI in global grasslands (Figs. 6 and 8), indicating that higher wetness offset the response of GPP and NEP to moderate grazing. This was likely because grazing-induced biomass reduction was less under wet condition in grasslands (Jin et al., 2023). However, not all grazing intensities aligned with the finding that the response of NEP increased with the increase in wetness index, likely due to severe loss of vegetation resilience under heavy grazing. Heavy grazing, even under higher wetness, reduced NEP and AGB in global grasslands (Fig. 2, 5 and 6C), resulting in a severe decline or even loss of their resilience, which could not be restored to control levels in long-term heavy grazing (Irob et al., 2023).

Schematic summary of the effects of grazing intensity on ecosystem carbon dioxide (CO2) fluxes and biomass in global grasslands in the meta-analysis.

The black and blue arrows indicate negative effects on overall grazing and grazing intensities, respectively. Asterisks (*) indicate significant effects on variables at P < 0.05. GPP, gross primary productivity; ER, ecosystem respiration; NEP, net ecosystem productivity; LG, light grazing; MG, moderate grazing; HG, heavy grazing.

Both our experiment and meta-analysis showed that aboveground biomass, grazing intensity and wetness index were key factors regulating the responses of ecosystem CO2 fluxes to grazing (Fig. S3, S8 and S9), with their responses intensifying with higher aboveground biomass and wetness index. Although grazing reduced C3 plant biomass, it had no significant effects on C4 plant biomass or the C3:C4 plant biomass ratio (Fig. S7). Furthermore, variation in plant community composition did not significantly explain the response of NEP to grazing (Fig. S8), indicating that the photosynthetic pathway composition of the plant community was not the primary driver mediating grazing effects on ecosystem carbon sequestration (Delgado-Balbuena et al., 2013). Grazing reduced aboveground biomass and plant respiration (Liu et al., 2020; Wang et al., 2025), thereby decreasing GPP and ER in both field experiment and meta-analysis (Fig. 4). Plants enhanced photosynthetic structures and physiological functions under grazing pressure. These adaptations improved CO2 assimilation and allocated more carbohydrates to compensate for the loss of aboveground biomass, thereby achieving compensatory growth (Wang et al., 2020; Yu et al., 2025). The increase in GPP and NEP in LG in 2020 was due to the increased relative growth rate in LG than in CK (Fig. S5), suggesting that grazing stimulated greater plant compensatory growth in years with more precipitation (Li et al., 2025). Moreover, appropriate grazing intensity facilitated the removal of senescent plant tissues, leading to increased leaf light capture capability and enhanced photosynthetic capacity (Shen et al., 2019). The aboveground and belowground biomass, as well as plant root/ shoot ratios were influenced by grazing activities such as feeding, trampling, and excretion (Cao et al., 2024). In the field experiment and global meta-analysis of this study, plant biomass (both aboveground and belowground biomass) was lower under heavy grazing (Figs. 2 and 5). Under heavy grazing conditions, where soil nutrients were often depleted, plants tended to prioritize biomass allocation to roots to enhance nutrient acquisition and utilization (Liu et al., 2024). Moreover, heavy grazing can stimulate C allocation to roots (Ren et al., 2024), which explains why the root/ shoot ratio was higher under heavy grazing (Figs. 2 and 5). In the wetter years (2020 and 2021), the aboveground biomass was higher while the root/ shoot ratio was lower than in the drier years. Heavy grazing affected the allocation strategy of aboveground and belowground biomass by plants.

Our long-term experiment and global meta-analysis jointly provided complementary evidence for understanding the effect of grazing on the ecosystem CO2 fluxes in grassland ecosystems. Long-term grazing field experiments revealed the mechanisms underlying the effects of grazing on ecosystem CO2 fluxes. Global meta-analysis further clarified the universality of these response patterns across different climatic zones and grassland types. It should be noticed that both field experiment and meta-analysis consistently revealed that wetness was an important factor in regulating the ecosystem CO2 fluxes in response to grazing. Under wetter conditions, light grazing promoted compensatory plant growth, enhanced leaf turnover and photosynthetic recovery capabilities, thereby maintaining or even enhancing the NEP (Morgan et al., 2016; Owensby et al., 2006). Conversely, under drier conditions or heavy grazing pressure, a decrease in aboveground biomass and weakened ecosystem resilience jointly limited the ecosystem carbon uptake capacity. These results suggested that wetness modulated the effects of grazing on NEP in grasslands. Moderate grazing may be sustainable under better water conditions, while heavy grazing may exceed the ecosystem resilience threshold even under relatively wet conditions, leading to sustained ecological degradation.

Limitations and implications for future study

Our results highlight that light grazing serves as a promising management strategy to promote ANPP and CO2 sequestration. However, some aspects still need to be further explored. The assessment of ecosystem CO2 fluxes was incomplete due to limited measurement time, area and sampling intervals. This may bias the representation of seasonal cycles and spatial heterogeneity, especially in the regions that were not monitored. Multiple time-point sampling should be adopted; the sampling frequency and points should also be increased in the future field study. Furthermore, key intermediate processes, such as soil nutrient availability, changes in community composition, and leaf photosynthetic physiological parameters, should also be investigated in the future. Our meta-analysis has limitations due to the relatively small sample size, which stems from the scarcity of studies across global grasslands exploring the effects of grazing intensity on ecosystem CO2 fluxes. The data of ecosystem CO2 fluxes from large-scale eddy covariance flux towers were not included in this meta-analysis, because there is an absence of a robust model to convert data between flux tower observations and chamber-based methods. In addition, future studies could conduct more experiments of ecosystem CO2 fluxes in South America, Africa and Oceania, which have large amounts of grasslands for grazing, but were rarely investigated. Furthermore, more drought or wetness indices should be investigated in future studies of grazing effects on ecosystem CO2 fluxes.

There are also limitations in evaluating the effects of grazing on grassland-livestock ecosystem CO2 fluxes in this study. Specifically, carbon emissions derived from animal respiration were not included in this study. Therefore, the results of the field experiment and meta-analysis in this study should not be interpreted as a complete carbon budget assessment of the grassland-livestock ecosystem. Our primary objective was to evaluate the carbon exchange processes between vegetation and soil under grazing disturbance. Thus, the grazing effects reported in the field experiment and meta-analysis of this study were responses at the vegetation-soil ecosystem scale rather than the net carbon balance of the entire grazing system.

In previous studies, meta-analysis literatures of effects of grazing intensities on NEP were limited and not adequate. Regardless of these limitations, the results of our field experiment and meta-analysis have significant implications for future studies and grassland production. The primary factors regulating ecosystem CO2 fluxes responses to grazing were identified as aboveground biomass, wetness index and grazing intensity. The NEP was positively affected by wetness index in LG, MG and HG. However, in both the typical steppe grazing experiment and the synthesis of global grasslands, the NEP resilience increased with higher wetness index specifically under light grazing, but not under heavy grazing. During wetter years, the NEP was even higher under LG compared to CK, whereas in drier years, no significant difference was observed between LG and CK (Fig. 2). This pattern aligns with the greater RGR and ANPP observed under elevated wetness index in light grazing treatment (Fig. S4). Globally, higher wetness index and aboveground biomass enhanced the response of NEP to grazing (Figs. 3, 7 and S14). However, even under higher wetness conditions, heavy grazing reduced NEP and AGB in global grasslands (Fig. 6), reflecting a severe decline, or even loss of resilience under long-term heavy grazing pressure. These results indicate that the negative impacts of long-term heavy grazing on NEP could not be offset by higher annual wetness index in global grasslands, underscoring the necessity of implementing strict grazing management measures. Additionally, moderate grazing under wetter years may be an appropriate management strategy to sustain grassland carbon balance while supporting livestock production.

Conclusion

In summary, the meta-analysis of this study presents the first comprehensive assessment of how annual wetness index affects the response of ecosystem CO2 fluxes to grazing across global grasslands. Our results showed that wetness index and aboveground biomass were key factors regulating the responses of ecosystem CO2 fluxes, and both exhibiting positive associations with these responses. Grazing reduced the response of GPP and ER in our field experiment of typical steppe and in the meta-analysis of global grasslands. Higher wetness indices offset the effect of moderate grazing rather than heavy grazing on NEP in global grasslands. These findings underscore the importance of developing region-specific grazing strategies guided by local wetness indices.